### 2024-10-27 docs/proposal/main.tex
add proposal template

%%%%%%%%%
% Generic article template using additional packages for table formatting
% New commands are defined for project participants: \katja, \flo, \dasha - 
% to draw attention to important remarks directly in the text
%%%%%%%%%

\documentclass{article}

\usepackage{graphicx}
\usepackage{multirow}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{adjustbox}

\newcommand{\katja}[1]{\textbf{\textcolor{teal}{(Katja) #1}}}
\newcommand{\dasha}[1]{\textbf{\textcolor{brown}{(Dasha) #1}}}
\newcommand{\flo}[1]{\textbf{\textcolor{red}{(Flo) #1}}}

\begin{document}

\title{Scientific Reasoning with LLMs}

\author{Group ??: Florian \and Ekaterina Kolos \and Daria }

\maketitle

\section{Introduction}

Scientific machine reasoning is ...

For our project, we intend to ...

Expected implementation details ...

\section{Background and Related Work}
% Large Language Models

Knowledge distillation allows to obtain smaller models capable of successfully following the behavior of larger "teacher" models. This is particularly useful for privacy reasons (running AI applications on mobile devices) and for cases where access to very large models in the cloud is not guaranteed (e.g. in cars). Some approaches to knowledge distillation include: \begin{itemize}
    \item distillation of outputs
    \item distillation of features on intermediate layers 
    \item CoT knowledge distillation \cite{shridhar2022distilling} 
    \item \katja{it's a draft TODO: add info from Dasha's slides}
\end{itemize}
Our goal is to test whether any of these techniques can be successfully used to obtain a model with scientific reasoning capabilities.

Agent-based architectures \cite{lin2024swiftsage} \cite{ghafarollahi2024sciagents}

Prompting techniques \cite{schulhoff2024prompt}

Scientific reasoning and the dataset \cite{lu2022learn}

\section{Methodology}
We postulate the following research questions: 
RQ1: Can smaller student LLMs be still effective at few-shot prompting for scientific reasoning?
RQ2: How does their performance compare with that of a teacher model?
RQ3: 
\paragraph{RQ1} 
We plan to

\paragraph{RQ2}
We plan to

\paragraph{RQ3}
We plan to

\paragraph{Metrics}

\section{Approximate Timeline}
Maybe just a paragraph here

\katja{Have reverted the doc to simple "article"}

\bibliography{mybibliography}
\bibliographystyle{plain}

\end{document}

### 2024-10-27 docs/proposal/main.tex
more edits on knowledge distillation

@@
 Some approaches to knowledge distillation include: \begin{itemize}
-    \item distillation of outputs
+    \item distillation of outputs, incl. CoT knowledge distillation, e.g. 
+    \begin{itemize}
+    \item self-consistency decoding: "The teacher model generates multiple CoT responses, and the student learns from the aggregate (self-consistent) reasoning paths, which can help the student model avoid common mistakes and output more consistent answers."
+    \item intermediate answers: "Generate CoT reasoning outputs from a large teacher model on reasoning.
+The student model will be fine-tuned using these CoT responses to produce intermediate reasoning steps."
+    \end{itemize}
     \item distillation of features on intermediate layers 
-    \item CoT knowledge distillation \cite{shridhar2022distilling} 
-    \item \katja{it's a draft TODO: add info from Dasha's slides}
+    %\item CoT knowledge distillation \cite{shridhar2022distilling} 
+    \item teacher model to produce training data for student model: "Use the teacher model to generate correct reasoning paths and analyze where the student model diverges.
+Distill additional targeted knowledge by emphasizing examples where the student struggles, focusing on error types like logical consistency or arithmetic errors."
+    \item architecture-based distillation
 \end{itemize}

### 2024-10-28 docs/proposal/main.tex
rewrite paragraph on knowledge distillation

--- 2024-10-27 docs/proposal/main.tex
@@
 
-Knowledge distillation allows to obtain smaller models capable of successfully following the behavior of larger "teacher" models.
+\paragraph{Knowledge distillation (KD)} allows to obtain smaller models capable of successfully following the behavior of larger teacher models.
 This is particularly useful for privacy reasons (running AI applications on mobile devices) and for cases where access to very large models in the cloud is not guaranteed (e.g.
 in cars).
-Some approaches to knowledge distillation include: \begin{itemize}
-    \item distillation of outputs
-    \item distillation of features on intermediate layers 
-    \item CoT knowledge distillation \cite{shridhar2022distilling} 
-    \item \katja{it's a draft TODO: add info from Dasha's slides}
-\end{itemize}
-Our goal is to test whether any of these techniques can be successfully used to obtain a model with scientific reasoning capabilities.
+The teacher model can be used to intelligently select examples on which the student model is trained (dataset distillation) \cite{yu2023dataset}, while \cite{li2024turning} rely on negative samples to show the student what incorrect answers or reasoning paths it should avoid to improve task accuracy.
+Training small models on a CoT reasoning path of a larger model was also shown to be a way to obtain a small student model replicating reasoning capabilities of teacher on downstream tasks \cite{magister2022teaching}, which is close to _response-based KD_ where the student model mimics the output of the teacher.
+More effective can be _feature-based KD_ where student also partially replicates the teacher on a feature-based level, when knowledge from specific layers of the teacher model is distilled into the student model \cite{sepahvand2022teacher}, while the student model's structure may be a quantized, simplified, or condensed version of the teacher's architecture \cite{gou2021knowledge}.
+
+\katja{Crazy idea: can we distill a reasoning path of a whole agent into a smaller model?
+I.e.
+instead of intelligent decision making that a large model performs for a complex task (search for X first, then look up Y, then formulate hypotheses Z) the student model learns the "protocols" of "customer support" and acts accordingly, without actually "understanding" why a particular action is good in a particular situation}
 

### 2024-10-28 docs/proposal/main.tex
add paragraph on prompting and rq idea

@@
 
-\author{Group ??: Florian \and Ekaterina Kolos \and Daria }
+%we'll probably have a group number later
+%names are given in alphabetical order
+\author{Stuttgart Team: Florian Dreyer \and Ekaterina Kolos \and Daria Matyash}
 
@@
 in cars).
-The teacher model can be used to intelligently select examples on which the student model is trained (dataset distillation) \cite{yu2023dataset}, while \cite{li2024turning} rely on negative samples to show the student what incorrect answers or reasoning paths it should avoid to improve task accuracy.
+The teacher model can be used to intelligently select examples on which the student model is trained (dataset distillation) \cite{yu2023dataset}, or provide on negative samples to show the student what incorrect answers or reasoning paths it should avoid to improve task accuracy  \cite{li2024turning}.
 Training small models on a CoT reasoning path of a larger model was also shown to be a way to obtain a small student model replicating reasoning capabilities of teacher on downstream tasks \cite{magister2022teaching}, which is close to _response-based KD_ where the student model mimics the output of the teacher.
@@
 
+\paragraph{Prompting techniques}.
+A prompt is a command a user gives to a generative LLM used to guide its output for a downstream task, which contains several of the following: a role (persona) the LLM has to follow (e.g.
+"You are a helpful assistant"), a precise task description, examples (exemplars) for analogous learning (few-shot prompting), requirements on the process of reasoning, style instructions, output format requirements, additional information for in-context learning, emotion prompting components (highlighting why the task or a particular requirement is important).
+Prompts can be combined into sequences or graphs (with complex branching and parallelism).
+Chain-of-thought prompts result in better reasoning, with variations including self-ask, \cite{schulhoff2024prompt}.
+Furthermore, prompts can be compressed and automatically optimized to improve efficiency and reduce costs \cite{chang2024efficient}.
+\cite{ge2023context} compress a long context 4x into memory slots; \cite{zhou2022large} make an LLM generate various prompting instructions that are then tested by another LLM on 24 NLP tasks to chose the best one.
+
+Hard prompts (discrete words of natural language) are opposed to soft prompts (learnable continuous representations).
+
 Agent-based architectures \cite{lin2024swiftsage} \cite{ghafarollahi2024sciagents}
-
-Prompting techniques \cite{schulhoff2024prompt}
+Reasoning and Acting (ReAct)
 
@@
 RQ2: How does their performance compare with that of a teacher model?
-RQ3: 
+RQ3: Can an LLM-agent's behavior be distilled into a single model?
 \paragraph{RQ1} 

### 2024-10-28 docs/proposal/main.tex
modify paragraph on prompting

@@
 
+\paragraph{Prompting techniques}.
+A prompt is a command a user gives to a generative LLM used to guide its output for a downstream task, which contains several of the following: a role (persona) the LLM has to follow (e.g.
+"You are a helpful assistant"), a precise task description, examples (exemplars) for analogous learning (few-shot prompting), requirements on the process of reasoning, style instructions, output format requirements, additional information for in-context learning, emotion prompting components (highlighting why the task or a particular requirement is important).
+Prompts can be combined into sequences or graphs (with complex branching and parallelism).
+Of particular interest are small requirements on the reasoning process that can significantly improve performance when added to the prompt, such as "Rephrase and expand the question, and respond" \cite{}, or "Let’s think step by step" \cite{kojima2022}.
+Chain-of-thought prompts improve reasoning capabilities by asking the model to speak its process of thinking out loud, with variations including self-ask, step-back prompting, thread-of-thought, least-to-most prompting (decomposing then solving), plan-and-solve prompting, tree-of-thought, recursion-of-thought and many more \cite{schulhoff2024prompt}.
+%with the resulting systems possibly using multiple CoT reasoning trajectories to select the most likely one (e.g.
+Uncertainty-Routed CoT Prompting, or making an LLM generate various prompting instructions that are then tested by another LLM on 24 NLP tasks to chose the best one \cite{zhou2022large}).
+%katja: I omit this to save space as not directly relevant
+Furthermore, prompts themselves can be compressed and automatically optimized to improve efficiency and reduce costs \cite{chang2024efficient}.
+\cite{ge2023context} compress a long context 4x into memory slots, while \cite{weston2023system} ask the LLM to first summarize the prompt and then execute it.
+Self-criticism and ensembling techniques can further be used to improve reasoning capabilities \cite{schulhoff2024prompt}.
+
+Hard prompts (discrete words of natural language) are opposed to soft prompts (learnable continuous representations).
+
 \paragraph{Knowledge distillation (KD)} allows to obtain smaller models capable of successfully following the behavior of larger teacher models.
@@
 in cars).
-The teacher model can be used to intelligently select examples on which the student model is trained (dataset distillation) \cite{yu2023dataset}, or provide on negative samples to show the student what incorrect answers or reasoning paths it should avoid to improve task accuracy  \cite{li2024turning}.
-Training small models on a CoT reasoning path of a larger model was also shown to be a way to obtain a small student model replicating reasoning capabilities of teacher on downstream tasks \cite{magister2022teaching}, which is close to _response-based KD_ where the student model mimics the output of the teacher.
-More effective can be _feature-based KD_ where student also partially replicates the teacher on a feature-based level, when knowledge from specific layers of the teacher model is distilled into the student model \cite{sepahvand2022teacher}, while the student model's structure may be a quantized, simplified, or condensed version of the teacher's architecture \cite{gou2021knowledge}.
+The teacher model can be used to intelligently select examples on which the student model is trained (dataset distillation) \cite{yu2023dataset}, or provide on negative samples to show the student what incorrect answers or reasoning paths it should avoid to improve task accuracy  \cite{li2024turning} (c.g.
+contrastive CoT).
+Training small models on a CoT reasoning path of a larger model was also shown to be a way to obtain a small student model replicating reasoning capabilities of teacher on downstream tasks \cite{magister2022teaching}, which is close to \textit{response-based KD} where the student model mimics the output of the teacher.
+More effective can be \textit{feature-based KD} where student also partially replicates the teacher on a feature-based level, when knowledge from specific layers of the teacher model is distilled into the student model \cite{sepahvand2022teacher}, while the student model's structure may be a quantized, simplified, or condensed version of the teacher's architecture \cite{gou2021knowledge}.
 
@@
 instead of intelligent decision making that a large model performs for a complex task (search for X first, then look up Y, then formulate hypotheses Z) the student model learns the "protocols" of "customer support" and acts accordingly, without actually "understanding" why a particular action is good in a particular situation}
-
-\paragraph{Prompting techniques}.
-A prompt is a command a user gives to a generative LLM used to guide its output for a downstream task, which contains several of the following: a role (persona) the LLM has to follow (e.g.
-"You are a helpful assistant"), a precise task description, examples (exemplars) for analogous learning (few-shot prompting), requirements on the process of reasoning, style instructions, output format requirements, additional information for in-context learning, emotion prompting components (highlighting why the task or a particular requirement is important).
-Prompts can be combined into sequences or graphs (with complex branching and parallelism).
-Chain-of-thought prompts result in better reasoning, with variations including self-ask, \cite{schulhoff2024prompt}.
-Furthermore, prompts can be compressed and automatically optimized to improve efficiency and reduce costs \cite{chang2024efficient}.
-\cite{ge2023context} compress a long context 4x into memory slots; \cite{zhou2022large} make an LLM generate various prompting instructions that are then tested by another LLM on 24 NLP tasks to chose the best one.
-
-Hard prompts (discrete words of natural language) are opposed to soft prompts (learnable continuous representations).
 

### 2024-10-29 docs/proposal/main.tex
Matrikelnummer
Update main.tex

@@
 %names are given in alphabetical order
-\author{Stuttgart Team: Florian Dreyer \and Ekaterina Kolos \and Daria Matyash}
+\author{Stuttgart Team: Florian Dreyer (3667877) \and Ekaterina Kolos \and Daria Matiash (3668740)}
 
@@
 
-Scientific machine reasoning is ...
+Scientific machine reasoning is the application of AI techniques to simulate scientific reasoning processes, such as  data interpretation, hypothesis generation, and causal inference.
+It facilitates scientific inquiry and discovery by enabling machines to independently analyze scientific data, generate hypotheses, and make predictions.
 

### 2024-10-29 docs/proposal/main.tex
Update main.tex
format template
Soft prompting

@@
 
-\documentclass{article}
+\documentclass[10pt]{article}
+\usepackage[a4paper, margin=1.25in]{geometry}
 
@@
 \usepackage{adjustbox}
+
+\usepackage{titlesec}
+\titlespacing*{\section}{0pt}{*0.8}{*0.8}
 
@@
 
-Hard prompts (discrete words of natural language) are opposed to soft prompts (learnable continuous representations).
+Soft prompting is an alternative to fine-tuning where the model is frozen while only a small number of "soft" prompt parameters are trained \cite{lester2021prompt}.
+These parameters are used to guide the model in the right direction.
 
@@
 \paragraph{Metrics}
+We'll evaluate model's performance with question answering accuracy domain-wise in order to have a fair comparison with leaderboards.
 

### 2024-10-29 docs/proposal/main.tex
Update main.tex

@@
 %names are given in alphabetical order
-\author{Stuttgart Team: Florian Dreyer (3667877) \and Ekaterina Kolos \and Daria Matiash (3668740)}
+\author{Stuttgart Team: Florian Dreyer (3667877) \and Ekaterina Kolos (3688474) \and Daria Matiash (3668740)}
 
@@
 We'll evaluate model's performance with question answering accuracy domain-wise in order to have a fair comparison with leaderboards.
+A few metrics used in machine translation such as BLEU-1, BLEU-4 and BERTScore can also be used, but the main metric we'll rely on is accuracy for question answering.
 

### 2024-10-29 docs/proposal/main.tex
Dataset
Timeline

@@
 
+\paragraph{Dataset}
+We plan to use the ScienceQA dataset. ScienceQA is a multimodal dataset containing science questions and their answers.
+
 \paragraph{Metrics}
@@
 \section{Approximate Timeline}
-Maybe just a paragraph here
+-Now - mid November: 
+  Preliminary baseline tests: small LLM relying only on commonsense reasoning / only on pretraining 
+  zero-shot, few-shot
+  optional: stronger baseline – same LLM with better prompting techniques like CoT/Self-Ask
+-mid November - mid December:
+  Construct agentic / RAG pipeline with “big smart” model + additional resources.
+  optional: soft-prompting for better workflow^ to distill from
+-mid December - mid January:
+  Knowledge distillation experiments.
 

### 2024-10-29 docs/proposal/main.tex
RS1&2

@@
 We postulate the following research questions: 
-RQ1: Can smaller student LLMs be still effective at few-shot prompting for scientific reasoning?
-RQ2: How does their performance compare with that of a teacher model?
+RQ1: Can we build a LLM Agent to improve the LLMs performance on science questions?
+RQ2: Does Soft Prompting improve the performance of the LLM (Agent)?
 RQ3: Can an LLM-agent's behavior be distilled into a single model?

### 2024-10-29 docs/proposal/main.tex
RQ3

@@
 \paragraph{RQ3}
-We plan to
+In order to distill the knowledge from the model fine-tuned in previous steps, we plan to do the model distillation with Chain-of-Thought Prompting for Reasoning approach.
+Following techniques described in \cite{magister2022teaching} \cite{wei2022chain}, the student model will be fine-tuned using these CoT responses to produce intermediate reasoning steps.
+At the same time, the teacher model generates multiple CoT responses, and the student learns from the aggregate (self-consistent) reasoning paths.
+The described approach can help the student model avoid common mistakes and output more consistent answers.
 

### 2024-10-29 docs/proposal/main.tex
Update main.tex
modify introduction

@@
 
-For our project, we intend to ...
+For our project, we intend to compare different approaches around generative large language models to generate solutions for school science problems.
+On an existing benchmark with human solutions, we are planning to compare: a pure LLM with zero-shot and few-shot prompting (and possibly some more fine-grained prompting techniques), an agentic approach with ability to query external resources like textbooks in physics, maths etc, and soft-prompting techniques.
+We then plan to attempt knowledge distillation to a smaller LLM.
 
@@
 Prompts can be combined into sequences or graphs (with complex branching and parallelism).
-Of particular interest are small requirements on the reasoning process that can significantly improve performance when added to the prompt, such as "Rephrase and expand the question, and respond" \cite{}, or "Let’s think step by step" \cite{kojima2022}.
+Of particular interest are small requirements on the reasoning process that can significantly improve performance when added to the prompt, such as
+%"Rephrase and expand the question, and respond" \cite{}, or 
+"Let’s think step by step" \cite{kojima2022large}. 
 Chain-of-thought prompts improve reasoning capabilities by asking the model to speak its process of thinking out loud, with variations including self-ask, step-back prompting, thread-of-thought, least-to-most prompting (decomposing then solving), plan-and-solve prompting, tree-of-thought, recursion-of-thought and many more \cite{schulhoff2024prompt}.
@@
 \paragraph{Dataset}
-We plan to use the ScienceQA dataset. ScienceQA is a multimodal dataset containing science questions and their answers.
+We plan to use the ScienceQA dataset \cite{lu2022learn}. ScienceQA is a multimodal dataset containing science questions and their answers.
 

### 2024-10-29 docs/proposal/main.tex
RQs

@@
 \paragraph{RQ2}
-We plan to
+We plan to use Soft Prompting on the LLM used to guide it towards better reasoning.
 
 \paragraph{RQ3}
-In order to distill the knowledge from the model fine-tuned in previous steps, we plan to do the model distillation with Chain-of-Thought Prompting for Reasoning approach.
+In order to distill the knowledge from the obtained model previous steps, we plan to do the model distillation with Chain-of-Thought Prompting for Reasoning approach.
 Following techniques described in \cite{magister2022teaching} \cite{wei2022chain}, the student model will be fine-tuned using these CoT responses to produce intermediate reasoning steps.

### 2024-10-29 docs/proposal/main.tex
update KD core idea

--- 2024-10-29 docs/proposal/main.tex
@@
 
-Expected implementation details ...
-
 \section{Background and Related Work}
@@
 
-\katja{Crazy idea: can we distill a reasoning path of a whole agent into a smaller model?
+This way, the student model learns the "protocols" of how to solve different problems, as well as the essence of the relevant knowledge selected by the teacher LLM. 
+%\katja{Crazy idea: can we distill a reasoning path of a whole agent into a smaller model?
 I.e.

### 2024-10-29 docs/proposal/main.tex
Update main.tex

@@
 \paragraph{Dataset}
-We plan to use the ScienceQA dataset \cite{lu2022learn}. ScienceQA is a multimodal dataset containing science questions and their answers.
+We plan to use the ScienceQA dataset \cite{lu2022learn}.
+The dataset includes a variety of science-related multimodal multiple-choice questions together with annotations of the answers that provide relevant lectures and explanations in Nature Science, Language Science and Social Science.
 

### 2024-10-29 docs/proposal/main.tex
RQ2
paragraph on agents

--- 2024-10-29 docs/proposal/main.tex
@@
 
+\paragraph{Agents}.
+An important step forward in using LLMs are agent-based architectures \cite{lin2024swiftsage} \cite{ghafarollahi2024sciagents}.
+They allow the LLM to act as an intelligent being capable of planning a solution of a complex task, while resorting to external resources on the way to acquire in-context the knowledge it does not have from pre-training.
+The retrieved information, results of invoking tools, such as calculators, and of interactions with the environment are appended to the prompt (or, in a more advanced scenario, to the summarized memory).
+Optionally, a reflexion step (explicit reasoning on all the accumulated information) is added before allowing further generation of the final response.
+
 \paragraph{Knowledge distillation (KD)} allows to obtain smaller models capable of successfully following the behavior of larger teacher models.
@@
 instead of intelligent decision making that a large model performs for a complex task (search for X first, then look up Y, then formulate hypotheses Z) the student model learns the "protocols" of "customer support" and acts accordingly, without actually "understanding" why a particular action is good in a particular situation}
-
-Agent-based architectures \cite{lin2024swiftsage} \cite{ghafarollahi2024sciagents}
-Reasoning and Acting (ReAct)
 

### 2024-10-29 docs/proposal/main.tex

@@
 We plan to use Soft Prompting on the LLM used to guide it towards better reasoning.
+To achieve this we will add learnable prompt parameters to the base LLM we use.
+While training these parameters the rest of the model will be frozen \cite{lester2021prompt}.
 

### 2024-10-29 docs/proposal/main.tex
RQ2
final version

@@
 \documentclass[10pt]{article}
-\usepackage[a4paper, margin=1.25in]{geometry}
+\usepackage[a4paper, margin=1.0in]{geometry}
 
@@
 
-\paragraph{Prompting techniques}.
-A prompt is a command a user gives to a generative LLM used to guide its output for a downstream task, which contains several of the following: a role (persona) the LLM has to follow (e.g.
+\paragraph{Prompting techniques} A prompt is a command a user gives to a generative LLM used to guide its output for a downstream task, which contains several of the following: a role (persona) the LLM has to follow (e.g.
 "You are a helpful assistant"), a precise task description, examples (exemplars) for analogous learning (few-shot prompting), requirements on the process of reasoning, style instructions, output format requirements, additional information for in-context learning, emotion prompting components (highlighting why the task or a particular requirement is important).
@@
 
-\paragraph{Agents}.
-An important step forward in using LLMs are agent-based architectures \cite{lin2024swiftsage} \cite{ghafarollahi2024sciagents}.
+\paragraph{Agents} An important step forward in using LLMs are agent-based architectures \cite{lin2024swiftsage} \cite{ghafarollahi2024sciagents}.
 They allow the LLM to act as an intelligent being capable of planning a solution of a complex task, while resorting to external resources on the way to acquire in-context the knowledge it does not have from pre-training.
@@
 
-Scientific reasoning and the dataset \cite{lu2022learn}
-
 \section{Methodology}
@@
 RQ3: Can an LLM-agent's behavior be distilled into a single model?
+
 \paragraph{RQ1} 
-We plan to
+We plan to start by prompting a multimodal model with reasoning capabilities with a zero-shot and few-shot prompting settings.
+A foundation model for this step could be a T4/T5 model or a multimodal Llama model.
+This simple generation will be compared with an agent-based approach using the same model, which will now include augmented retrieval of information on the scientific task from domain-specific texts.
 
@@
 We'll evaluate model's performance with question answering accuracy domain-wise in order to have a fair comparison with leaderboards.
-A few metrics used in machine translation such as BLEU-1, BLEU-4 and BERTScore can also be used, but the main metric we'll rely on is accuracy for question answering.
+A few metrics used in machine translation such as BLEU-1, BLEU-4 and BERTScore can also be used to evaluate the reasoning steps.
+%, but the main metric we'll rely on is accuracy for question answering.
 
 \section{Approximate Timeline}
--Now - mid November: 
-  Preliminary baseline tests: small LLM relying only on commonsense reasoning / only on pretraining 
-  zero-shot, few-shot
-  optional: stronger baseline – same LLM with better prompting techniques like CoT/Self-Ask
--mid November - mid December:
-  Construct agentic / RAG pipeline with “big smart” model + additional resources.
-  optional: soft-prompting for better workflow^ to distill from
--mid December - mid January:
-  Knowledge distillation experiments.
-
-\katja{Have reverted the doc to simple "article"}
+\begin{enumerate}
+	\item \textbf{Now - mid November}: 
+		\begin{itemize}
+  			\item Preliminary baseline tests: small LLM relying only on commonsense reasoning / only on pretraining (zero-shot, few-shot).
+			\item Optional: stronger baseline – same LLM with better prompting techniques like CoT/Self-Ask.
+		\end{itemize}
+	\item \textbf{mid November - mid December}:
+		\begin{itemize}
+  			\item Construct agentic / RAG pipeline with “big smart” model + additional resources.
+  			\item Optional: soft-prompting for better workflow to distill from.
+		\end{itemize}
+	\item \textbf{mid December - mid January}:
+  		\begin{itemize}
+			\item Knowledge distillation experiments.
+		\end{itemize}
+\end{enumerate}
 

### 2024-10-29 docs/proposal/main.tex
fix bibliography & prepare pdf
add specific foundation models to proposal

@@
 \documentclass[10pt]{article}
-\usepackage[a4paper, margin=1.0in]{geometry}
+\usepackage[a4paper, margin=.90in]{geometry}
 
@@
 
+\bibliography{main}
+\usepackage[style=numeric, backend=biber]{biblatex}
+
 \begin{document}
@@
 It facilitates scientific inquiry and discovery by enabling machines to independently analyze scientific data, generate hypotheses, and make predictions.
-
 For our project, we intend to compare different approaches around generative large language models to generate solutions for school science problems.
@@
 %katja: I omit this to save space as not directly relevant
-Furthermore, prompts themselves can be compressed and automatically optimized to improve efficiency and reduce costs \cite{chang2024efficient}.
+%Furthermore, prompts themselves can be compressed and automatically optimized to improve efficiency and reduce costs \cite{chang2024efficient}.
 \cite{ge2023context} compress a long context 4x into memory slots, while \cite{weston2023system} ask the LLM to first summarize the prompt and then execute it.
 Self-criticism and ensembling techniques can further be used to improve reasoning capabilities \cite{schulhoff2024prompt}.
-
 Soft prompting is an alternative to fine-tuning where the model is frozen while only a small number of "soft" prompt parameters are trained \cite{lester2021prompt}.
@@
 They allow the LLM to act as an intelligent being capable of planning a solution of a complex task, while resorting to external resources on the way to acquire in-context the knowledge it does not have from pre-training.
-The retrieved information, results of invoking tools, such as calculators, and of interactions with the environment are appended to the prompt (or, in a more advanced scenario, to the summarized memory).
+The retrieved information, results of invoking tools, such as calculators, and of interactions with the environment are appended to the prompt (or, in a more advanced scenario, to the summarized memory \cite{ge2023context}).
 Optionally, a reflexion step (explicit reasoning on all the accumulated information) is added before allowing further generation of the final response.
@@
 \paragraph{Knowledge distillation (KD)} allows to obtain smaller models capable of successfully following the behavior of larger teacher models.
-This is particularly useful for privacy reasons (running AI applications on mobile devices) and for cases where access to very large models in the cloud is not guaranteed (e.g.
+This is particularly useful for privacy reasons (mobile AI apps) and when access to large models in the cloud is not guaranteed (e.g.
 in cars).
-The teacher model can be used to intelligently select examples on which the student model is trained (dataset distillation) \cite{yu2023dataset}, or provide on negative samples to show the student what incorrect answers or reasoning paths it should avoid to improve task accuracy  \cite{li2024turning} (c.g.
+The teacher model can be used to intelligently select examples on which the student model is trained (dataset distillation) \cite{yu2023dataset}, or provide negative samples to show the student what incorrect answers or reasoning paths it should avoid to improve task accuracy \cite{li2024turning} (c.f.
 contrastive CoT).
 Training small models on a CoT reasoning path of a larger model was also shown to be a way to obtain a small student model replicating reasoning capabilities of teacher on downstream tasks \cite{magister2022teaching}, which is close to \textit{response-based KD} where the student model mimics the output of the teacher.
-More effective can be \textit{feature-based KD} where student also partially replicates the teacher on a feature-based level, when knowledge from specific layers of the teacher model is distilled into the student model \cite{sepahvand2022teacher}, while the student model's structure may be a quantized, simplified, or condensed version of the teacher's architecture \cite{gou2021knowledge}.
+Alternatively, with \textit{feature-based KD} the knowledge from specific layers of the teacher model is distilled into the student model \cite{sepahvand2022teacher}, while the student model's structure may be a quantized, simplified, or condensed version of the teacher's architecture \cite{gou2021knowledge}.
 
-This way, the student model learns the "protocols" of how to solve different problems, as well as the essence of the relevant knowledge selected by the teacher LLM. 
+In our project, we plan to develop an extended version of CoT knowledge distillation, where both the reasoning paths and the essence of the external knowledge the teacher agent has retrieved for the tasks are acting as training data for the student model.
+This way the teacher agent "defines" the "protocols" of solving complex tasks, while the student model tries to "follow" the learnt "instructions" and memorize important information for similar tasks.
+If we succeed, such small models could be a valuable asset in practice in enterprise environments, to encode important protocols and parts of project documentation, replacing expensive intelligent agents for specific tasks.
+
 %\katja{Crazy idea: can we distill a reasoning path of a whole agent into a smaller model?
@@
 We postulate the following research questions: 
-RQ1: Can we build a LLM Agent to improve the LLMs performance on science questions?
-RQ2: Does Soft Prompting improve the performance of the LLM (Agent)?
-RQ3: Can an LLM-agent's behavior be distilled into a single model?
+RQ1: Can we build a LLM agent to improve the LLMs performance on science problems?
+RQ2: Does Soft Prompting improve the performance of the LLM (agent)?
+RQ3: Can an LLM agent's behavior be distilled into a single model?
 
@@
 We plan to start by prompting a multimodal model with reasoning capabilities with a zero-shot and few-shot prompting settings.
-A foundation model for this step could be a T4/T5 model or a multimodal Llama model.
-This simple generation will be compared with an agent-based approach using the same model, which will now include augmented retrieval of information on the scientific task from domain-specific texts.
+Preliminary list of foundation models to be compared as baselines is: CLIP, GPT4o, T5, BigGAN, Gemini; small models: TinyCLIP, T5-Small; Llama models.
+This simple generation will be compared with an agent-based approach using the same models, which will now include augmented retrieval of information on the scientific task from domain-specific texts.
 
 \paragraph{RQ2}
-We plan to use Soft Prompting on the LLM used to guide it towards better reasoning.
+We plan to use Soft Prompting as one of the parameter-efficient fine-tuning techniques on the LLM used to guide it towards better reasoning.
 To achieve this we will add learnable prompt parameters to the base LLM we use.
@@
 \paragraph{RQ3}
-In order to distill the knowledge from the obtained model previous steps, we plan to do the model distillation with Chain-of-Thought Prompting for Reasoning approach.
-Following techniques described in \cite{magister2022teaching} \cite{wei2022chain}, the student model will be fine-tuned using these CoT responses to produce intermediate reasoning steps.
+In order to distill the knowledge from the system obtained at previous steps, we plan to do the model distillation with Chain-of-Thought Prompting for Reasoning approach.
+Following techniques described in \cite{magister2022teaching} \cite{wei2022chain}, the student model will be fine-tuned using these CoT responses together with any additional information the agent has retrieved from external sources to produce intermediate reasoning steps.
 At the same time, the teacher model generates multiple CoT responses, and the student learns from the aggregate (self-consistent) reasoning paths.
-The described approach can help the student model avoid common mistakes and output more consistent answers.
+The foundation model to serve as student LLM should be small, multimodal, capable of reasoning and capable of acquiring new distilled knowledge.
+A preliminary list of candidate models: small Llama models, MiniGPT-4, small FLAVA models distilUNITER, lxmert, distilled ViLT, to be also compared with text-only DeBERTa-V3 Small / Base, DistilBERT, GPT-2 Medium.
+%The described approach can help the student model avoid common mistakes and output more consistent answers.
 
@@
 We'll evaluate model's performance with question answering accuracy domain-wise in order to have a fair comparison with leaderboards.
-A few metrics used in machine translation such as BLEU-1, BLEU-4 and BERTScore can also be used to evaluate the reasoning steps.
+The reasoning steps can be evaluated with semantic similarity metrics (e.g.
+adopted from machine translation) such as BLEU, METEOR, ROUGE, BLANC, and BERTScore and more.
 %, but the main metric we'll rely on is accuracy for question answering.
@@
 		\begin{itemize}
-  			\item Preliminary baseline tests: small LLM relying only on commonsense reasoning / only on pretraining (zero-shot, few-shot).
-			\item Optional: stronger baseline – same LLM with better prompting techniques like CoT/Self-Ask.
+  			\item Preliminary baseline tests: compare "pure LLM"s relying only on commonsense reasoning / only on pretraining (zero-shot, few-shot).
+			\item Stronger baseline – same LLM with better prompting techniques like CoT/Self-Ask.
 		\end{itemize}
@@
 
-\bibliography{mybibliography}
-\bibliographystyle{plain}
+\printbibliography
 

### 2024-10-29 docs/proposal/main.tex
kept only multimodal FMs in preliminary list

@@
 We plan to start by prompting a multimodal model with reasoning capabilities with a zero-shot and few-shot prompting settings.
-Preliminary list of foundation models to be compared as baselines is: CLIP, GPT4o, T5, BigGAN, Gemini; small models: TinyCLIP, T5-Small; Llama models.
+Preliminary list of foundation models to be compared as baselines is: CLIP, GPT4o (GPT4-V), T5, BigGAN, Gemini; small models: TinyCLIP, T5-Small; Llama models.
 This simple generation will be compared with an agent-based approach using the same models, which will now include augmented retrieval of information on the scientific task from domain-specific texts.
@@
 The foundation model to serve as student LLM should be small, multimodal, capable of reasoning and capable of acquiring new distilled knowledge.
-A preliminary list of candidate models: small Llama models, MiniGPT-4, small FLAVA models distilUNITER, lxmert, distilled ViLT, to be also compared with text-only DeBERTa-V3 Small / Base, DistilBERT, GPT-2 Medium.
+A preliminary list of candidate models: small Llama and LLaVA models, MiniGPT-4, small FLAVA models, distilUNITER, lxmert, distilled ViLT.
 %The described approach can help the student model avoid common mistakes and output more consistent answers.

### 2024-11-11 helpers.py
Added function for creating the prompt

@@ -1,0 +1,17 @@
+def build_prompt(question_data, use_lecture=False, use_solution=False):
+    question = base_prompt.get_question_text(question_data)
+    choices = base_prompt.get_choice_text(question_data, [choice_num for choice_num in range(5)])
+    hint = base_prompt.get_context_text(question_data, False)
+    image = question_data['image']
+    task = question_data['task']
+    input_prompt = f'Question: {question}\n Task: {task}\n Choices: {choices}\n Hint: {hint}'
+    if use_lecture:
+        lecture = f'\n Lecture: {question_data["lecture"]}'
+        input_prompt += lecture
+    if use_solution and question_data["solution"]:
+        solution = f'\n Solution: {question_data["solution"]}'
+        input_prompt += solution
+    prompt = [input_prompt]
+    if image:
+        prompt.append(image)
+    return prompt

### 2024-11-11 docs/paper/main.tex
start paper draft from thesis proposal

--- 2024-10-29 docs/proposal/main.tex
@@
 
-\paragraph{Dataset}
-We plan to use the ScienceQA dataset \cite{lu2022learn}.
+\section{Dataset}
+This study is based on the ScienceQA dataset \cite{lu2022learn}.
 The dataset includes a variety of science-related multimodal multiple-choice questions together with annotations of the answers that provide relevant lectures and explanations in Nature Science, Language Science and Social Science.
+
+Data points are structured as follows:
+\begin{verbatim}
+{'image': <PIL.PngImagePlugin.PngImageFile image mode=RGB size=200x202>,
+ 'question': 'Is diorite a mineral or a rock?',
+ 'choices': ['rock', 'mineral'],
+ 'answer': 0,
+ 'hint': 'Diorite has the following properties:\nno fixed crystal structure\nnaturally occurring\nnot a pure substance\ncoarse-grained texture\nsolid\nnot made by organisms',
+ 'task': 'closed choice',
+ 'grade': 'grade8',
+ 'subject': 'natural science',
+ 'topic': 'earth-science',
+ 'category': 'Rocks and minerals',
+ 'skill': 'Identify rocks and minerals',
+ 'lecture': 'Minerals are the building blocks of rocks.
+A rock can be made of one or more minerals.\nMinerals and rocks have the following properties:\nProperty | Mineral | Rock\nIt is a solid.
+| Yes | Yes\nIt is formed in nature.
+| Yes | Yes\nIt is not made by organisms.
+| Yes | Yes\nIt is a pure substance.
+| Yes | No\nIt has a fixed crystal structure.
+| Yes | No\nYou can use these properties to tell whether a substance is a mineral, a rock, or neither.\nLook closely at the last three properties:\nMinerals and rocks are not made by organisms.\nOrganisms make their own body parts.
+For example, snails and clams make their shells.
+Because they are made by organisms, body parts cannot be  minerals or rocks.\nHumans are organisms too.
+So, substances that humans make by hand or in factories are not minerals or rocks.\nA mineral is a pure substance, but a rock is not.\nA pure substance is made of only one type of matter.
+Minerals are pure substances, but rocks are not.
+Instead, all rocks are mixtures.\nA mineral has a fixed crystal structure, but a rock does not.\nThe crystal structure of a substance tells you how the atoms or molecules in the substance are arranged.
+Different types of minerals have different crystal structures, but all minerals have a fixed crystal structure.
+This means that the atoms and molecules in different pieces of the same type of mineral are always arranged the same way.\nHowever, rocks do not have a fixed crystal structure.
+So, the arrangement of atoms or molecules in different pieces of the same type of rock may be different!',
+ 'solution': 'The properties of diorite match the properties of a rock. So, diorite is a rock.'}
+\end{verbatim}
+
+For some datapoints, the image data is missing, while the answer can be deduced from text alone (e.g.
+questions on 'subject': 'language science', 'topic': 'figurative-language', 'skill': 'Interpret figures of speech' (e.g.
+irony recognition).
+
+For other datapoints, however, the image data can be missing, although required to make an answer, e.g.
+\begin{verbatim}
+{'image': None,
+ 'question': 'What information supports the conclusion that Rick inherited this trait?',
+ 'choices': ["Rick's coworker also has curly hair.",
+  "Rick's biological father has curly hair.",
+  'Rick and his biological parents have brown hair.'],
+ 'answer': 1,
+ 'hint': 'Read the description of a trait.\nRick has curly hair.',
+ 'task': 'closed choice',
+ 'grade': 'grade7',
+ 'subject': 'natural science',
+ 'topic': 'biology',
+ 'category': 'Genes to traits',
+ 'skill': 'Inherited and acquired traits: use evidence to support a statement',
+ 'lecture': "Organisms, including people, have both inherited and acquired traits.
+Inherited and acquired traits are gained in different ways.\nInherited traits are passed down from biological parents to their offspring through genes.
+Genes are pieces of hereditary material that contain the instructions that affect inherited traits.
+Offspring receive their genes, and therefore gain their inherited traits, from their biological parents.
+Inherited traits do not need to be learned.\nAcquired traits are gained during a person's life.
+Some acquired traits, such as riding a bicycle, are gained by learning.
+Other acquired traits, such as scars, are caused by the environment.
+Parents do not pass acquired traits down to their offspring.",
+ 'solution': ''}
+\end{verbatim}
+
+Apart from the image field, missing values can occur in lecture and solution fields. The statistics in table \ref{table:dataset-stats} give an overview on the amount of problematic datapoints.
+
+\begin{table}[h!]
+\centering
+\begin{tabular}{ll}\toprule
+	missing & frequency \\\midrule
+	solution & 9.3\% \\
+	lecture & 15.7\% \\
+	solution and/or lecture & 24.2\% \\
+	image & 50.6\%\\\bottomrule
+\end{tabular}
+\caption{Defect datapoints in dataset}
+\label{table:dataset-stats}
+\end{table}
+
+\paragraph{Retained data} As 50.6\% of the datapoints are missing an image, we decide to process both text-only and text+image datapoints similarly with multimodal models, attaching the image to the prompt if it is available.
+For now, we will only work with datapoints with non-empty solution AND with non-empty lecture. All the other datapoints are disregarded.
 
@@
 %, but the main metric we'll rely on is accuracy for question answering.
+
+\section{Experiments}
+\subsection{Zero-shot and Few-shot Benchmarking}
+We start by zero-shot benchmarking existing multimodal models in three simple prompt settings, partly repeating with the official evaluation suite:
+\begin{enumerate}
+	\item \textbf{question - choices - hint - image - task}, with the following prompt:
+
+	\item \textbf{question - choices - hint - image - task + lecture}, with the following prompt:
+
+	\item \textbf{question - choices - hint - image - task + lecture + solution}, with the following prompt:
+
+\end{enumerate}
+
+For the first two the model has to output \textbf{answer - solution}, for the third - \textbf{answer} (the easiest set up meant to ensure the model can deduce the correct answer from an already provided correct solution).
+
+We evaluate: OpenAI, Gemini, MistralAI, as well as small local models. %TODO: list models.
+
+The metrics used for evaluation at this point are: accuracy, \texttt{BLEU-1}, \texttt{BLEU-4}, \texttt{ROUGE}, similarity.
 

### 2024-11-11 docs/paper/main.tex
docs: add prompts of benchmarking experiments

@@
 	\item \textbf{question - choices - hint - image - task}, with the following prompt:
-
+	\texttt{f'Question: {question}\n Task: {task}\n Choices: {choices}\n Hint: {hint}'}
 	\item \textbf{question - choices - hint - image - task + lecture}, with the following prompt:
-
+	\texttt{f'Question: {question}\n Task: {task}\n Choices: {choices}\n Hint: {hint}\n Lecture: {lecture}'}
 	\item \textbf{question - choices - hint - image - task + lecture + solution}, with the following prompt:
-
+	\texttt{f'Question: {question}\n Task: {task}\n Choices: {choices}\n Hint: {hint}\n Lecture: {lecture}\n Solution: {solution}'}	
 \end{enumerate}

### 2024-11-13 evaluate_acc.py
evaluation_files

@@ -1,0 +1,74 @@
+import os
+import json
+import argparse
+import warnings
+import pandas as pd
+
+warnings.filterwarnings('ignore')
+
+
+def get_acc_with_contion(res_pd, key, values):
+    if isinstance(values, list):
+        total_pd = res_pd[res_pd[key].isin(values)]
+    else:
+        total_pd = res_pd[res_pd[key] == values]
+    correct_pd = total_pd[total_pd['true_false'] == True]
+    if len(total_pd) == 0:
+        return -1
+    acc = "{:.2f}".format(len(correct_pd) / len(total_pd) * 100)
+    return acc
+
+
+def get_scores(res_pd):
+    num = res_pd.shape[0]
+    for index, row in res_pd.iterrows():
+        print(index)
+        res_pd.loc[index, 'no_context'] = True if (not row['hint'] and not row['image']) else False
+        res_pd.loc[index, 'has_text'] = True if row['hint'] else False
+        res_pd.loc[index, 'has_image'] = True if row['image'] else False
+        res_pd.loc[index, 'has_text_image'] = True if (row['hint'] and row['image']) else False
+
+        label = row['answer']
+        pred = row['answer_output']
+        res_pd.loc[index, 'pred'] = pred
+        res_pd.loc[index, 'true_false'] = (label == pred)
+
+    # accuracy scores
+    acc_average = len(res_pd[res_pd['true_false'] == True]) / num * 100
+    #assert result_file.split('_')[-1] == "{:.3f}.json".format(acc_average)
+
+    scores = {
+        'acc_natural':
+        get_acc_with_contion(res_pd, 'subject', 'natural science'),
+        'acc_social':
+        get_acc_with_contion(res_pd, 'subject', 'social science'),
+        'acc_language':
+        get_acc_with_contion(res_pd, 'subject', 'language science'),
+        'acc_has_text':
+        get_acc_with_contion(res_pd, 'has_text', True),
+        'acc_has_image':
+        get_acc_with_contion(res_pd, 'has_image', True),
+        'acc_no_context':
+        get_acc_with_contion(res_pd, 'no_context', True),
+        'acc_grade_1_6':
+        get_acc_with_contion(res_pd, 'grade', ['grade1', 'grade2', 'grade3', 'grade4', 'grade5', 'grade6']),
+        'acc_grade_7_12':
[…]

### 2024-11-13 evaluation.py

@@ -1,0 +1,48 @@
+import pandas as pd
+import json, ast, re
+from datasets import load_dataset
+from sentence_transformers import SentenceTransformer
+from evaluations import caculate_bleu, caculate_rouge, caculate_similariry
+from evaluate_acc import *
+
+
+def eval(RES_DIR:str, MODEL_NAME:str):
+    for i in range(1,5):
+        df = pd.read_csv(f"{RES_DIR}/{MODEL_NAME}_val_output_setting_{i}.csv", sep="\t")
+        df = pd.read_excel("../res/see.xlsx")
+        answer_pred_col, solution_pred_col = "answer_output", "solution_output"
+        df[[answer_pred_col, solution_pred_col]] = df["output"].apply(string2json).apply(pd.Series)
+        answer_preds = df[answer_pred_col].tolist()
+        solution_preds =  df[solution_pred_col].tolist()
+        scores = get_scores(df)
+        print_scores(scores)
+        metrics = calculate_metrics_solutions(solution_preds , df)
+        scores.update(metrics)
+        df_save = pd.DataFrame(scores)
+        df_save.to_csv(f"{RES_DIR}/{MODEL_NAME}_val_metrics.csv", sep="\t", encoding="utf-8")
+       
+
+def string2json(text):
+    # convert json-like string to "answer" and "solution"
+    dict_json = ast.literal_eval(re.sub("```|\n|json", "", text))
+    return dict_json["answer"], dict_json["solution"]
+
+
+def calculate_metrics_solutions(results, data):
+    ## BLEU
+    bleu1 = caculate_bleu(results, data, gram=1)
+    bleu4 = caculate_bleu(results, data, gram=4)
+    print("BLEU-1: %.3f" % bleu1)
+    print("BLEU-4: %.3f" % bleu4)
+
+    ## Rouge-L
+    rouge = caculate_rouge(results, data)
+    print("ROUGE-L: %.3f" % rouge)
+
+    ## Similarity
+    model = SentenceTransformer('sentence-transformers/all-MiniLM-L6-v2').cuda()
+    model = SentenceTransformer('sentence-transformers/all-MiniLM-L6-v2')
+    similarity = caculate_similariry(results, data, model)
+    print("Similariry: %.3f" % similarity)
+
+    return {"BLEU-1": [bleu1], "BLEU-4": [bleu4], "ROUGE": [rouge], "similarity": similarity}

### 2024-11-13 evaluations.py

@@ -1,0 +1,102 @@
+import re
+from rouge import Rouge
+from nltk.translate.bleu_score import sentence_bleu
+from sentence_transformers import util
+
+
+def extract_explanation(text):
+    text = re.sub(r"The answer is [A-Z]. BECAUSE: ", "", text)
+    return text
+
+
+########################
+## BLEU
+########################
+def tokenize(text):
+    tokens = re.split(r'\s|\.', text)
+    tokens = [t for t in tokens if len(t) > 0]
+    return tokens
+
+
+def bleu_score(reference, hypothesis, gram):
+    reference_tokens = tokenize(reference)
+    hypothesis_tokens = tokenize(hypothesis)
+
+    if gram == 1:
+        bleu = sentence_bleu([reference_tokens], hypothesis_tokens, (1., ))  # BELU-1
+    elif gram == 2:
+        bleu = sentence_bleu([reference_tokens], hypothesis_tokens, (1. / 2., 1. / 2.))  # BELU-2
+    elif gram == 3:
+        bleu = sentence_bleu([reference_tokens], hypothesis_tokens, (1. / 3., 1. / 3., 1. / 3.))  # BELU-3
+    elif gram == 4:
+        bleu = sentence_bleu([reference_tokens], hypothesis_tokens, (1. / 4., 1. / 4., 1. / 4., 1. / 4.))  # BELU-4
+
+    return bleu
+
+
+def caculate_bleu(results, df, gram):
+    bleus = []
+    for qid, output in enumerate(results):
+        prediction = extract_explanation(output)
+        target = df["lecture"][qid] + " " + df["solution"][qid]
+        target = target.strip()
+        if target == "":
+            continue
+        bleu = bleu_score(target, prediction, gram)
+        bleus.append(bleu)
+
+    avg_bleu = sum(bleus) / len(bleus)
+
+    return avg_bleu
+
+
+########################
+## Rouge-L
+########################
+def score_rouge(str1, str2):
+    rouge = Rouge(metrics=["rouge-l"])
+    scores = rouge.get_scores(str1, str2, avg=True)
+    rouge_l = scores['rouge-l']['f']
+    return rouge_l
+
+
+def caculate_rouge(results, df):
+    rouges = []
+    for qid, output in enumerate(results):
+        prediction = extract_explanation(output)
+        target = df["lecture"][qid] + " " + df["solution"][qid]
+        target = target.strip()
+        if prediction == "":
+            continue
+        if target == "":
+            continue
+        rouge = score_rouge(target, prediction)
[…]

### 2024-11-14 evaluate_acc.py
Update evaluate_acc.py

@@ -20,7 +20,6 @@ def get_acc_with_contion(res_pd, key, values):
 
 
 def get_scores(res_pd):
-    num = res_pd.shape[0]
     for index, row in res_pd.iterrows():
         print(index)
         res_pd.loc[index, 'no_context'] = True if (not row['hint'] and not row['image']) else False

### 2024-11-14 evaluate_acc.py
Update evaluate_acc.py

@@ -20,8 +20,8 @@ def get_acc_with_contion(res_pd, key, values):
 
 
 def get_scores(res_pd):
+    num = res_pd.shape[0]
     for index, row in res_pd.iterrows():
-        print(index)
         res_pd.loc[index, 'no_context'] = True if (not row['hint'] and not row['image']) else False
         res_pd.loc[index, 'has_text'] = True if row['hint'] else False
         res_pd.loc[index, 'has_image'] = True if row['image'] else False

### 2024-11-14 evaluation.py
Update evaluation.py

@@ -12,6 +12,7 @@ def eval(RES_DIR:str, MODEL_NAME:str):
         df = pd.read_excel("../res/see.xlsx")
         answer_pred_col, solution_pred_col = "answer_output", "solution_output"
         df[[answer_pred_col, solution_pred_col]] = df["output"].apply(string2json).apply(pd.Series)
+        df[answer_pred_col] = df[answer_pred_col].apply(lambda x: int(x))
         answer_preds = df[answer_pred_col].tolist()
         solution_preds =  df[solution_pred_col].tolist()
         scores = get_scores(df)

### 2024-11-14 evaluation.py
Update evaluation.py

@@ -19,8 +19,10 @@ def eval(RES_DIR:str, MODEL_NAME:str):
         print_scores(scores)
         metrics = calculate_metrics_solutions(solution_preds , df)
         scores.update(metrics)
-        df_save = pd.DataFrame(scores)
-        df_save.to_csv(f"{RES_DIR}/{MODEL_NAME}_val_metrics.csv", sep="\t", encoding="utf-8")
+        dict_save = {'model': MODEL_NAME, 'setting': SETTING_NUM}
+        dict_save.update(scores)
+        df_save = pd.DataFrame(dict_save)
+        df_save.to_csv(f"{RES_DIR}/{MODEL_NAME}_val_metrics_setting_{i}.csv", sep="\t", encoding="utf-8")
        
 
 def string2json(text):

### 2024-11-17 docs/paper/main.tex
llava 1.5 update results and metrics

--- 2024-11-11 docs/paper/main.tex
@@
 
+Finally, hint can be missing, e.g.
+the text-only task 367 on subject language science: "What is the source of allusion in the sentence below?
+Edward picked up his pace on the trail as his spidey sense began to tingle" with choices ["Italian history", "a comic book"], or the text-only task 3 on the topic of capitalization: "Which correctly shows the title of a play?" with choices ["A breath of Fresh Air", "A Breath of Fresh Air"].
+
 \begin{table}[h!]
@@
 \paragraph{Retained data} As 50.6\% of the datapoints are missing an image, we decide to process both text-only and text+image datapoints similarly with multimodal models, attaching the image to the prompt if it is available.
-For now, we will only work with datapoints with non-empty solution AND with non-empty lecture. All the other datapoints are disregarded.
+For now, we will only work with datapoints with non-empty solution AND with non-empty lecture. All the other datapoints are disregarded,
+%\begin{verbatim}
+%filtered\_validation\_data = validation\_data.filter(lambda example: example["lecture"] != "" and example["solution"] != "")
+%\end{verbatim}
+which results in \textbf{3216} datapoints kept in validation data.
 
@@
 	\item \textbf{question - choices - hint - image - task}, with the following prompt:
-
+	\begin{verbatim}{f'Question: {question}\n Task: {task}\n Choices: {choices}\n Hint: {hint}'}\end{verbatim}
 	\item \textbf{question - choices - hint - image - task + lecture}, with the following prompt:
-
+	\begin{verbatim}{f'Question: {question}\n Task: {task}\n Choices: {choices}\n Hint: {hint}\n 
+	Lecture: {lecture}'}\end{verbatim}
 	\item \textbf{question - choices - hint - image - task + lecture + solution}, with the following prompt:
-
+	\begin{verbatim}{f'Question: {question}\n Task: {task}\n Choices: {choices}\n Hint: {hint}\n 
+	Lecture: {lecture}\n Solution: {solution}'}\end{verbatim}
 \end{enumerate}
@@
 We evaluate: OpenAI, Gemini, MistralAI, as well as small local models. %TODO: list models.
+
+\subsubsection{Local Models from HuggingFace}
+
+\paragraph{Pre-experiments} We were not successful with processing with small non-chat models.
+For the given text + image, the models would return short non-meaningful answers, like "Yes" to a "X or Y" question, or "no idea" to a completely defined task with all relevant context included.
+These were multimodal models below 1B on Huggingface, not tuned to be used in a chat scenario (and prepending a "You are a helpful assistant..." to the beginning of the prompt was not enough to fix this).
+We are not sure how to use those models for our purposes as of now.
+
+\paragraph{MiniGPT-4} This is a small multimodal model designed from ?? \cite{zhu2023minigpt}. It offers to build the model based on Llama or Vicuna. 
+\textbf{Unsuccessful experiment}: We tried to resort to an unofficial implementation: \texttt{wangrongsheng/MiniGPT-4-LLaMA-7B}. Loading the model was not successful (no preprocessor\_config.json). 
+Another issue: using BlipForConditionalGeneration and BlipProcessor: "You are using a model of type llama to instantiate a model of type blip.
+This is not supported for all configurations of models and can yield errors.".
+Model size: ~10GB.
+Processor: ~3.6GB.
+
+\paragraph{LLaVa} The exact model used is: \texttt{llava-hf/llava-1.5-7b-hf} (and there is a larger sister 13b model).
+It is based on LLaMA and (?) Vicuna fine-tuned on GPT-generated multimodal instruction-following data.
+It expects inputs as conversation, allowing the preprocessor to convert textual input into format "USER: {request} ASSISTANT: " where the assistant has to fill in the rest. 
+We encountered the following issues while applying this model version on ScienceQA data: 
+\begin{itemize}
+	\item although prompted like larger models, this model returned answer strings and not choice index; this may additionally raise an issue of the returned answer not being entirely part of the choice list;
+	\item although prompted to generate JSON, the model would sometimes generate a different string, e.g.
+simply repeating part of the task.
+Unexpected format happened in ??
+\% of all answers.
+We disregarded these answers, considering them all wrong, which results in a bit more pessimistic estimation of the model's performance.
+The metrics on the correctly formatted answers alone are also given.
+	\item the model currently does not accept inputs without an image (text-only); according to a huggingface discussion\footnote{https://huggingface.co/llava-hf/llava-1.5-7b-hf/discussions/38}, the authors claim to have accidentally removed this possibility in a recent update.
+We worked around this by producing placeholder empty 10x10 white images and passing them as part of the input when the image in the dataset was missing.
+We also appended the suffix "Ignore image" to the end of the prompt, because the model would otherwise try to hallucinate a relation between the text and the image (e.g.
+"Why is the sky blue" + placeholder image results in "The sky is blue on the picture because ...");
+	\item the model does not really seem to support batch processing; although officially the technical possibility exists, it has been reported online \footnote{https://huggingface.co/llava-hf/llava-1.5-13b-hf/discussions/10} that the model would only take into account the first image in the batch, which makes the processing useless.
+We did not risk to attempt the supported or custom batching and processed the filtered validation dataset 1 datapoint by 1, which took 10+ hours on GPU for inference.
+\end{itemize}
+
+There is another available model LLaVANext, which recently started to support batching, and should have better performance. 
+%Issues: https://huggingface.co/llava-hf/llava-v1.6-mistral-7b-hf/discussions/34
+
+Note: I later (after running the model) understood that the original model is this one: `liuhaotian/llava-v1.5-7b`.
+The results may thus be different from what would be expected from the model's paper.
+
+\paragraph{BLIP-2}
+
+\paragraph{LLaMA models}
+
+\paragraph{Gemma2}
+
+\paragraph{Qwen models} e.g.  Qwen2.5-Coder
+
+\paragraph{}
+
+\begin{table}
+\centering
+\begin{tabular}{rcc}
+	\toprule
+	Model name & llava-1.5-7b-hf & \\
+	Size & 7B & \\
+	Published & 09.2023 & \\
+	Type & SLM & \\
+	Chat model & yes & \\
+	Parent models & LLaMA, Vicuna & \\
+	Data & GPT-generated multimodal instruction-following data & \\
+	\bottomrule
+\end{tabular}
+\caption{Model Comparison: Benchmarked Models}
+\end{table}
 

### 2024-11-17 foundation_models_small_multi_(2).py

@@ -1,0 +1,325 @@
+# -*- coding: utf-8 -*-
+"""Foundation Models Small Multi (2)
+
+Automatically generated by Colab.
+
+Original file is located at
+    https://colab.research.google.com/drive/1ynqpZK_YPcbB_H_4kiXEEXaLWR9VY5FB
+
+# Load Data
+"""
+
+!pip install datasets
+
+from datasets import load_dataset
+dataset = load_dataset("derek-thomas/ScienceQA")
+validation_data = dataset["validation"]
+validation_data
+
+"""### This is a normal datapoint"""
+
+validation_data[3500]
+
+validation_data[3500]["image"]
+
+"""### This is an abnormal datapoint with no image
+But the text is enough to figure out the solution
+"""
+
+validation_data[0]
+
+"""### This is a bad abnormal datapoint
+Image is part of the task input and is missing. The solution is also missing.
+"""
+
+validation_data[4240]
+
+filtered_validation_data = validation_data.filter(lambda example: example["lecture"] != "" and example["solution"] != "")
+filtered_validation_data
+
+"""# Experiment with Models"""
+
+import torch
+
+device = torch.device("cuda" if torch.cuda.is_available() else "cpu")
+device
+
+from abc import ABC, abstractmethod
+
+class Model(ABC):
+    """Abstract base class for models with custom processing."""
+
+    def __init__(self, model_name: str):
+        self.model_name = model_name
+        self.model = self.load_model()
+        self.processor = self.load_processor()
+
+    @abstractmethod
+    def load_model(self):
+        """Load the model architecture and weights."""
+        pass
+
+    @abstractmethod
+    def load_processor(self):
+        """Load the processor associated with the model."""
+        pass
+
+    @abstractmethod
+    def process(self, *args, **kwargs):
+        """Define the processing logic for the model."""
+        pass
+
+"""## Preliminary Experiments"""
+
+class LLaVAModel(Model):
+    def load_model(self):
+        from transformers import LlavaForConditionalGeneration
+        return LlavaForConditionalGeneration.from_pretrained(
+            self.model_name,
+            torch_dtype=torch.float16,
+            low_cpu_mem_usage=True)
+
+    def load_processor(self):
+        from transformers import AutoProcessor
+        return AutoProcessor.from_pretrained(self.model_name)
+
+    def process(self, text_input, image, **kwargs):
[…]

### 2024-11-17 foundation_models_small_multi_(2).py
cleanup

@@ -1,325 +1,0 @@
-# -*- coding: utf-8 -*-
-"""Foundation Models Small Multi (2)
-
-Automatically generated by Colab.
-
-Original file is located at
-    https://colab.research.google.com/drive/1ynqpZK_YPcbB_H_4kiXEEXaLWR9VY5FB
-
-# Load Data
-"""
-
-!pip install datasets
-
-from datasets import load_dataset
-dataset = load_dataset("derek-thomas/ScienceQA")
-validation_data = dataset["validation"]
-validation_data
-
-"""### This is a normal datapoint"""
-
-validation_data[3500]
-
-validation_data[3500]["image"]
-
-"""### This is an abnormal datapoint with no image
-But the text is enough to figure out the solution
-"""
-
-validation_data[0]
-
-"""### This is a bad abnormal datapoint
-Image is part of the task input and is missing. The solution is also missing.
-"""
-
-validation_data[4240]
-
-filtered_validation_data = validation_data.filter(lambda example: example["lecture"] != "" and example["solution"] != "")
-filtered_validation_data
-
-"""# Experiment with Models"""
-
-import torch
-
-device = torch.device("cuda" if torch.cuda.is_available() else "cpu")
-device
-
-from abc import ABC, abstractmethod
-
-class Model(ABC):
-    """Abstract base class for models with custom processing."""
-
-    def __init__(self, model_name: str):
-        self.model_name = model_name
-        self.model = self.load_model()
-        self.processor = self.load_processor()
-
-    @abstractmethod
-    def load_model(self):
-        """Load the model architecture and weights."""
-        pass
-
-    @abstractmethod
-    def load_processor(self):
-        """Load the processor associated with the model."""
-        pass
-
-    @abstractmethod
-    def process(self, *args, **kwargs):
-        """Define the processing logic for the model."""
-        pass
-
-"""## Preliminary Experiments"""
-
-class LLaVAModel(Model):
-    def load_model(self):
-        from transformers import LlavaForConditionalGeneration
-        return LlavaForConditionalGeneration.from_pretrained(
-            self.model_name,
-            torch_dtype=torch.float16,
-            low_cpu_mem_usage=True)
-
-    def load_processor(self):
-        from transformers import AutoProcessor
-        return AutoProcessor.from_pretrained(self.model_name)
-
-    def process(self, text_input, image, **kwargs):
[…]

### 2024-11-21 docs/paper/main.tex
update

@@
 \section{Experiments}
+
 \subsection{Zero-shot and Few-shot Benchmarking}
@@
 We evaluate: OpenAI, Gemini, MistralAI, as well as small local models. %TODO: list models.
+
+\section{Benchmark}
+% description of metrics, why they were used + analytics of results
+\subsection{Accuracy}
+The highest performance was observed in settings that included solutions, indicating that the models were generally capable of extracting relevant information effectively.
+Among these, settings combining both lecture and solution information typically yielded slightly better results than those with solutions alone.
+`GPT-4` consistently received highest ratings, especially in scenarios in which lecture and/or solution information was included.
+The Gemini family of models showed better robustness in many scenarios and came in second overall.
+Notably, Gemini models achieved an average accuracy advantage of 5\% over GPT-family models in the "pure" task context (without lecture or solution information).
+The `LLaVAModel1.5-7b`, on the other hand, demonstrated the lowest average accuracy scores, particularly in environments without lecture and solution data.
+
+\subsection{Cosine similarity}
+\subsection{BLEU-1}
+Across settings, higher BLEU-1 scores were observed in settings 3 and 4 (those incorporating solutions and/or lecture information), indicating that models performed better with richer contextual data.
+The `GPT-4o-mini` model demonstrated superior performance, surpassing `GPT-4`, in achieving the highest results.
+Lower scores in settings 1 and 2 suggest that models, particularly `LLaVAModel1.5-7b`, may have had trouble correctly aligning their outputs with reference words when given less auxiliary information, as indicated by lower scores in settings 1 and 2.
+
+\subsection{BLEU-4}
+Scores were consistently higher in settings 3 and 4, particularly for  models, demonstrating their ability to produce more contextually precise outputs when both lecture and solution information were provided.
+Gemini models had higher scores in comparison to other models, which represents a better capability of information extraction.
+`LlaVAModel`, at the same time, gained the highest score in the setting without solutions, which indicates the ability of extracting knowledge from massive textual data.
+
+\subsection{ROUGE}
+The highest ROUGE scores were observed in setting 3, where both lectures and solutions were available, highlighting the importance of comprehensive input for producing informative responses.
+Setting 1 showed lower ROUGE scores across all models, reflecting limited informativeness when models were provided with minimal context.
+`LlaVAModel` anf `GPT-4` had highest ROUGE-scores on settings without solutions.
+Outputs of `GPT-4` without knowledge of the correct answer had higher ROUGE score than `GPT-4o-mini` given solutions in the input.
+
+\subsection{Cosine similarity}
+Semantic alignment with the reference is assessed by the similarity metric.
+Gemini and its variants outperformed other models, demonstrating their great capacity to capture the semantic structure of enriched input.
+Similarity scores were highest in settings 3 and 4.
+LLaVAModel1.5-7b, on the other hand, received the lowest score in setting 1, indicating that it mostly depends on lecture or solution data to generate outputs that are semantically accurate.
+This result emphasises how more data affects the resulting content's semantic integrity.
+`GPT-4o-mini` and `LLaVAModel` performed better than other models in environments lacking solution information, indicating that they can produce reliable results even with limited  contextual guidance.
+
+\subsection{Summary}
+Overall, the findings demonstrate how important enriched contextual data, such as lectures and solutions, in enhancing model performance across all metrics.
+The GPT-4o-mini model outperformed GPT-4 in multiple instances, indicating the efficacy of smaller, more targeted designs, even if GPT-4 continuously received excellent accuracy and informativeness scores.
+While LLaVAModel1.5-7b failed without extra input, highlighting its dependence on extensive contextual knowledge to function well, the Gemini family of models demonstrated remarkable robustness and semantic alignment, particularly in enriched situations.
+Lecture information doesn’t really affect models’ performance, it can even decrease the ratio of correct answers.
 

### 2024-11-28 docs/paper/main.tex
update

@@
 adopted from machine translation) such as BLEU, METEOR, ROUGE, BLANC, and BERTScore and more.
+\subsection{Multiple choice Evaluation}
+Owing to the simplicity of the test format, only accuracy score is computed, following original evaluation strategy by \cite{lu2022learn}. This metric is considered of the highest importantce 
+\subsection{Answer Reasoning Evaluation}
+Due to the pecularity of scientific texts and approaches to the evaluation of automatically generated texts, the following evaluation approaches were chosen.
+Our validation process is consisted of the metrics introduced in the original paper \cite{lu2022learn}  with a few additional metrics and techniques that we consider of high importance in regards to the scientific textual data.
+\subsubsubsection{BLEU}
+
+\subsubsubsection{ROUGE}
+\subsubsubsection{METEOR}
+
 %, but the main metric we'll rely on is accuracy for question answering.

### 2024-12-07 helpers.py
First steps knowledge distillation

@@ -1,7 +1,30 @@
+def get_question_text(problem):
+    question = problem['question']
+    return question
+
+
+def get_choice_text(probelm, options):
+    choices = probelm['choices']
+    choice_list = []
+    for i, c in enumerate(choices):
+        choice_list.append("({}) {}".format(options[i], c))
+    choice_txt = " ".join(choice_list)
+    return choice_txt
+
+
+def get_context_text(problem, use_caption):
+    txt_context = problem['hint']
+    img_context = problem['caption'] if use_caption else ""
+    context = " ".join([txt_context, img_context]).strip()
+    if context == "":
+        context = "N/A"
+    return context
+
+
 def build_prompt(question_data, use_lecture=False, use_solution=False):
-    question = base_prompt.get_question_text(question_data)
-    choices = base_prompt.get_choice_text(question_data, [choice_num for choice_num in range(5)])
-    hint = base_prompt.get_context_text(question_data, False)
+    question = get_question_text(question_data)
+    choices = get_choice_text(question_data, [choice_num for choice_num in range(5)])
+    hint = get_context_text(question_data, False)
     image = question_data['image']
     task = question_data['task']
     input_prompt = f'Question: {question}\n Task: {task}\n Choices: {choices}\n Hint: {hint}'

### 2024-12-07 soft_prompting.py

@@ -17,7 +17,7 @@ class SoftPrompting:
             random_range: float = 0.5,
             **kwargs,
     ):
-        device = torch.device("cuda:0" if torch.cuda.is_available() else "cpu")
+        device = torch.device("cuda" if torch.cuda.is_available() else "cpu")
         model.to(device)
         for param in model.parameters():
             param.requires_grad = False

### 2024-12-09 docs/paper/main.tex
upd

@@
 \paragraph{Metrics}
-We'll evaluate model's performance with question answering accuracy domain-wise in order to have a fair comparison with leaderboards.
+We evaluate model's performance with question answering accuracy domain-wise in order to have a fair comparison with leaderboards.
 The reasoning steps can be evaluated with semantic similarity metrics (e.g.
-adopted from machine translation) such as BLEU, METEOR, ROUGE, BLANC, and BERTScore and more.
+adopted from machine translation) such as BLEU, METEOR, ROUGE, and cosine similarity and more.
 \subsection{Multiple choice Evaluation}
-Owing to the simplicity of the test format, only accuracy score is computed, following original evaluation strategy by \cite{lu2022learn}. This metric is considered of the highest importantce 
+Owing to the simplicity of the test format, only accuracy score is computed, following original evaluation strategy by \cite{lu2022learn}.
+It was determined that the computation of the overall score should assign the highest weight (\textbf{50\%}) to the ability to provide correct answers (accuracy score), as this represents the most critical aspect of performance evaluation.
 \subsection{Answer Reasoning Evaluation}
@@
 Our validation process is consisted of the metrics introduced in the original paper \cite{lu2022learn}  with a few additional metrics and techniques that we consider of high importance in regards to the scientific textual data.
-\subsubsubsection{BLEU}
 
-\subsubsubsection{ROUGE}
-\subsubsubsection{METEOR}
+\subsubsection{BLEU}
+BLEU-score (\cite{papineni2002bleu}) can measure how closely the model's generated explanation aligns with the human-authored explanation.
+BLEU-1 measures if the model uses the right scientific terms or key words (e.g., "photosynthesis," "temperature").
+However, it ignores word order and logical progression, so it can’t evaluate reasoning or explanation quality.
+That is why BLEU-4 score is also computed in order to capture both vocabulary and the arrangement of words into meaningful phrases, evaluating fluency and coherence.
+Scientific explanations can often convey the same idea with different wording, making semantic similarity hard to capture.
+BLEU might penalize valid but rephrased or concise answers and it does not account for deeper meaning or logical correctness, which are crucial in reasoning tasks.
+Since this metric can give valuable insights on the generated outputs, we do not consider these scores as the most important and both BLEU-1 and BLEU-4 scores have \textbf{5\%} impact.
+
+\subsubsection{ROUGE}
+ROUGE-L score, following evaluation strategy in \cite{lu2022learn}, helps analyze the content recall and sequence preservation of model-generated explanations.
+They highlight whether a model captures key scientific concepts and produces explanations that align with reference reasoning.
+However, ROUGE should be used in conjunction with other metrics to fully evaluate the logical and factual quality of generated reasoning.
+ROUGE-L will highlight the mismatch in logical flow, even though some content overlaps in reference and candidate sentences (\ref{example:1}, \ref{example:2}).
+It lead to the conclusion that this metric would not drastically affect the overall score and contribute (\textbf{5\%}).
+\begin{example}
+\label{example:1} 
+\begin{itemize}
+    \item \textbf{Reference:} Water evaporates, cools, and condenses to form clouds.
+    \item \textbf{Candidate:} Water cools, evaporates, and forms clouds.
+\end{itemize}
+\end{example}
+\begin{example}
+\label{example:2}
+\begin{itemize}
+    \item \textbf{Reference:} Plants absorb sunlight to produce energy.
+    \item \textbf{Candidate:} Sunlight is absorbed by plants to create energy.
+\end{itemize}
+\end{example}
+\subsubsubsection{Cosine similarity with Sentence Transformers}
+Cosine similarity using SentenceTransformer embeddings provides a robust method for evaluating model outputs in the ScienceQA dataset, focusing on the semantic alignment of generated explanations with references.
+It captures nuanced relationships and conceptual understanding, which gives more understanding of the generated outputs in comparison to ROUGE and BLEU scores and that is why it has more importance in final score computation.
+However, due to domain specificity it can miss some intricate cases like still considering high similarity of two sentences, even though it the reference can contain crucial information about carbon dioxide, while in the candidate sentence the answer was given in relation to oxygen reproduction.
+\subsubsection{METEOR}
+METEOR (\cite{banerjee2005meteor}) evaluates text similarity based on precision, recall, and alignment of semantically meaningful components like synonyms and stems, making it well-suited for capturing the nuanced language used in scientific reasoning.
+By incorporating a weighted F-score and penalties for excessive mismatches, METEOR provides a balanced evaluation of fluency and accuracy, crucial for assessing LLM-generated answers on the Science QA Dataset.
+Due to these balanced text evaluation, this metric has \textbf{10\%} impact on the overall score.
+% \subsubsection{Cosine similarity with SciBERT}
+% We extended the evaluation of cosine similarity using SciBERT (\cite{beltagy2019scibert}) embeddings.
+These vectors are more consistent with the domain-specific terminology and reasoning patterns present in the Science QA Dataset because it has been properly pre-trained on scientific literature.
+% SciBERT generates more accurate similarity embeddings because it better reflects the complex relationships between scientific concepts than general-purpose models like those in Sentence Transformers.
+% When assessing the logical and factual coherence of scientific reasoning, the structural and contextual semantics specific to scientific texts are reflected in the embeddings produced by SciBERT.
+% Assessments for jobs requiring sophisticated scientific understanding are more reliable when SciBERT is used for cosine similarity since it guarantees that the evaluation metric is domain-specific.
+
+\subsubsection{Cosine similarity with Sentence Transformers}
+Sentence Transformer models are optimized for sentence-level embeddings, which capture the overall contextual similarity of answers more effectively, making them suitable for general-purpose evaluation tasks like scientific reasoning.
+Unlike SciBERT, which is tailored to scientific text and could have been used as sentence embedder, Sentence Transformers are pre-trained on diverse datasets, enabling them to generalize better to variations in phrasing and style present in the Science QA Dataset.
+Their ability to handle broader linguistic nuances ensures that the evaluation accounts for both semantic correctness and the logical flow of responses, which is essential when assessing reasoning across a wide range of scientific topics.
+That is the reason for using Sentence Transformer model for measurements of cosine similarity between the lecture material and annotated explanations and models' outputs with impact weight of \textbf{15\%}.
 

### 2024-12-09 soft_prompting.py
Soft prompting for decoder only

@@ -18,7 +18,6 @@ class SoftPrompting:
             **kwargs,
     ):
         device = torch.device("cuda" if torch.cuda.is_available() else "cpu")
-        model.to(device)
         for param in model.parameters():
             param.requires_grad = False
 
@@ -34,37 +33,23 @@ class SoftPrompting:
 
         return model
 
-    def set_soft_prompt_embeds(
-            self,
-            soft_prompt_path: str,
-    ) -> None:
-        """
-        Args:
-            soft_prompt_path: torch soft prompt file path
-
-        """
-        device = torch.device("cuda:0" if torch.cuda.is_available() else "cpu")
-        self.soft_prompt = torch.load(
-            soft_prompt_path, map_location=device
-        )
+    def set_soft_prompt_embeds(self, soft_prompt_path: str) -> None:
+        device = torch.device("cuda" if torch.cuda.is_available() else "cpu")
+        self.soft_prompt = torch.load(soft_prompt_path, map_location=device)
         self.n_tokens = self.soft_prompt.num_embeddings
         print(f"Set soft prompt! (n_tokens: {self.n_tokens})")
 
     def initialize_soft_prompt(
-            self,
-            n_tokens: int = 20,
-            initialize_from_vocab: bool = True,
-            random_range: float = 0.5,
+            self, n_tokens: int = 20, initialize_from_vocab: bool = True, random_range: float = 0.5
     ) -> None:
         self.n_tokens = n_tokens
         if initialize_from_vocab:
             init_prompt_value = self.transformer.wte.weight[:n_tokens].clone().detach()
         else:
-            init_prompt_value = torch.FloatTensor(2, 10).uniform_(
+            init_prompt_value = torch.FloatTensor(n_tokens, self.config.n_embd).uniform_(
                 -random_range, random_range
             )
         self.soft_prompt = nn.Embedding(n_tokens, self.config.n_embd)
-        # Initialize weight
         self.soft_prompt.weight = nn.parameter.Parameter(init_prompt_value)
 
     def _cat_learned_embedding_to_input(self, input_ids) -> torch.Tensor:
@@ -73,8 +58,9 @@ class SoftPrompting:
         if len(list(inputs_embeds.shape)) == 2:
             inputs_embeds = inputs_embeds.unsqueeze(0)
 
-        # [batch_size, n_tokens, n_embd]
[…]

### 2024-12-09 docs/paper/main.tex
docs: paper ICML template

--- 2024-11-21 docs/paper/main.tex
@@
-%%%%%%%%%
-% Generic article template using additional packages for table formatting
-% New commands are defined for project participants: \katja, \flo, \dasha - 
-% to draw attention to important remarks directly in the text
-%%%%%%%%%
+%%%%%%%% ICML 2021 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%
 
-\documentclass[10pt]{article}
-\usepackage[a4paper, margin=.90in]{geometry}
+\documentclass{article}
 
+% Recommended, but optional, packages for figures and better typesetting:
+\usepackage{microtype}
 \usepackage{graphicx}
-\usepackage{multirow}
+\usepackage{subfigure}
+\usepackage{booktabs} % for professional tables
+
+% hyperref makes hyperlinks in the resulting PDF.
+% If your build breaks (sometimes temporarily if a hyperlink spans a page)
+% please comment out the following usepackage line and replace
+% \usepackage{icml2021} with \usepackage[nohyperref]{icml2021} above.
 \usepackage{hyperref}
-\usepackage{xcolor}
-\usepackage{booktabs}
-\usepackage{adjustbox}
 
-\usepackage{titlesec}
-\titlespacing*{\section}{0pt}{*0.8}{*0.8}
+% Attempt to make hyperref and algorithmic work together better:
+\newcommand{\theHalgorithm}{\arabic{algorithm}}
 
-\newcommand{\katja}[1]{\textbf{\textcolor{teal}{(Katja) #1}}}
-\newcommand{\dasha}[1]{\textbf{\textcolor{brown}{(Dasha) #1}}}
-\newcommand{\flo}[1]{\textbf{\textcolor{red}{(Flo) #1}}}
+% Use the following line for the initial blind version submitted for review:
+%\usepackage{icml2021}
 
-\bibliography{main}
-\usepackage[style=numeric, backend=biber]{biblatex}
+% If accepted, instead use the following line for the camera-ready submission:
+\usepackage[accepted]{icml2021}
 
+% The \icmltitle you define below is probably too long as a header.
+% Therefore, a short form for the running title is supplied here:
+\icmltitlerunning{Submission and Formatting Instructions for ICML 2021}
 \begin{document}
 
-\title{Scientific Reasoning with LLMs}
+\twocolumn[
+\icmltitle{Scientific Reasoning: \\
+           Assessment of Multimodal Generative LLMs}
 
-%we'll probably have a group number later
-%names are given in alphabetical order
-\author{Stuttgart Team: Florian Dreyer (3667877) \and Ekaterina Kolos (3688474) \and Daria Matiash (3668740)}
+% It is OKAY to include author information, even for blind
+% submissions: the style file will automatically remove it for you
+% unless you've provided the [accepted] option to the icml2021
+% package.
 
-\maketitle
+% List of affiliations: The first argument should be a (short)
+% identifier you will use later to specify author affiliations
+% Academic affiliations should list Department, University, City, Region, Country
+% Industry affiliations should list Company, City, Region, Country
+
+% You can specify symbols, otherwise they are numbered in order.
+% Ideally, you should not use this facility. Affiliations will be numbered
+% in order of appearance and this is the preferred way.
+\icmlsetsymbol{equal}{*}
+
+\begin{icmlauthorlist}
+\icmlauthor{Florian Dreyer (3667877)}{equal,cs}
+\icmlauthor{Ekaterina Kolos (3688474)}{equal,ims}
+\icmlauthor{Daria Matiash (3668740)}{equal,ims}
+\end{icmlauthorlist}
+
+\icmlaffiliation{cs}{Institut fuer KI}
+\icmlaffiliation{ims}{Institut fuer Maschinelle Sprachverarbeitung}
+
+\icmlcorrespondingauthor{Florian Dreyer}{c.vvvvv@googol.com}
+%\icmlcorrespondingauthor{Eee Pppp}{ep@eden.co.uk}
+
+% You may provide any keywords that you
+% find helpful for describing your paper; these are used to populate
+% the "keywords" metadata in the PDF but will not be shown in the document
+\icmlkeywords{Scientific Reasoning, VQA}
+
+\vskip 0.3in
+]
+% this must go after the closing bracket ] following \twocolumn[ ...
+
+% This command actually creates the footnote in the first column
+% listing the affiliations and the copyright notice.
+% The command takes one argument, which is text to display at the start of the footnote.
+% The \icmlEqualContribution command is standard text for equal contribution.
+% Remove it (just {}) if you do not need this facility.
+
+%\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
+\printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.
+
+\begin{abstract}
+This project assesses the capabilities of pre-trained multimodal LLMs to perform scientific reasoning tasks on multimodal data.
+We further explore how prompt tuning (soft prompting) and prompt engineering techniques can improve performance.
+We then attempt to distill knowledge to a small LLM.
+\end{abstract}
 
 \section{Introduction}
+\label{introduction}
 
+\textbf{SECTION NEEDS UPDATE}
 Scientific machine reasoning is the application of AI techniques to simulate scientific reasoning processes, such as  data interpretation, hypothesis generation, and causal inference.
@@
 
+\section{Motivation}
+\label{motivation}
+\textbf{TODO}
+
 \section{Background and Related Work}
+\label{related-work}
+
 % Large Language Models
@@
 We postulate the following research questions: 
-RQ1: Can we build a LLM agent to improve the LLMs performance on science problems?
-RQ2: Does Soft Prompting improve the performance of the LLM (agent)?
-RQ3: Can an LLM agent's behavior be distilled into a single model?
+%RQ1: Can we build a LLM agent to improve the LLMs performance on science problems?
+%RQ2: Does Soft Prompting improve the performance of the LLM (agent)?
+%RQ3: Can an LLM agent's behavior be distilled into a single model?
+RQ1: Does Soft Prompting improve the performance of the LLM for scientific reasoning?
+RQ2: Can a teacher model's knowledge be distilled into a student model's soft prompt?
 
-\paragraph{RQ1} 
-We plan to start by prompting a multimodal model with reasoning capabilities with a zero-shot and few-shot prompting settings.
+%\paragraph{RQ1} 
+%We plan to start by prompting a multimodal model with reasoning capabilities with a zero-shot and few-shot prompting settings.
 Preliminary list of foundation models to be compared as baselines is: CLIP, GPT4o (GPT4-V), T5, BigGAN, Gemini; small models: TinyCLIP, T5-Small; Llama models.
@@
 
-\paragraph{RQ2}
-We plan to use Soft Prompting as one of the parameter-efficient fine-tuning techniques on the LLM used to guide it towards better reasoning.
+%\paragraph{RQ2}
+%We plan to use Soft Prompting as one of the parameter-efficient fine-tuning techniques on the LLM used to guide it towards better reasoning.
 To achieve this we will add learnable prompt parameters to the base LLM we use.
 While training these parameters the rest of the model will be frozen \cite{lester2021prompt}.
+\paragraph{RQ1}
+Our soft prompting approach relies on prefix soft prompts as introduced in [TODO: cite Li 2021], that can be used both for encoder-decoder and for autoregressive architectures.
+The models we test here are autoregressive.
 
-\paragraph{RQ3}
-In order to distill the knowledge from the system obtained at previous steps, we plan to do the model distillation with Chain-of-Thought Prompting for Reasoning approach.
+%\paragraph{RQ3}
+%In order to distill the knowledge from the system obtained at previous steps, we plan to do the model distillation with Chain-of-Thought Prompting for Reasoning approach.
 Following techniques described in \cite{magister2022teaching} \cite{wei2022chain}, the student model will be fine-tuned using these CoT responses together with any additional information the agent has retrieved from external sources to produce intermediate reasoning steps.
 At the same time, the teacher model generates multiple CoT responses, and the student learns from the aggregate (self-consistent) reasoning paths.
-The foundation model to serve as student LLM should be small, multimodal, capable of reasoning and capable of acquiring new distilled knowledge.
+%The foundation model to serve as student LLM should be small, multimodal, capable of reasoning and capable of acquiring new distilled knowledge.
 A preliminary list of candidate models: small Llama and LLaVA models, MiniGPT-4, small FLAVA models, distilUNITER, lxmert, distilled ViLT.
@@
 \section{Dataset}
+\label{dataset}
 This study is based on the ScienceQA dataset \cite{lu2022learn}.
@@
 
-Data points are structured as follows:
-\begin{verbatim}
-{'image': <PIL.PngImagePlugin.PngImageFile image mode=RGB size=200x202>,
- 'question': 'Is diorite a mineral or a rock?',
- 'choices': ['rock', 'mineral'],
- 'answer': 0,
- 'hint': 'Diorite has the following properties:\nno fixed crystal structure\nnaturally occurring\nnot a pure substance\ncoarse-grained texture\nsolid\nnot made by organisms',
- 'task': 'closed choice',
- 'grade': 'grade8',
- 'subject': 'natural science',
- 'topic': 'earth-science',
- 'category': 'Rocks and minerals',
- 'skill': 'Identify rocks and minerals',
- 'lecture': 'Minerals are the building blocks of rocks.
+%Data points are structured as follows:
+%\begin{verbatim}
+%{'image': <PIL.PngImagePlugin.PngImageFile image mode=RGB size=200x202>,
+% 'question': 'Is diorite a mineral or a rock?',
+% 'choices': ['rock', 'mineral'],
+% 'answer': 0,
+% 'hint': 'Diorite has the following properties:\nno fixed crystal structure\nnaturally occurring\nnot a pure substance\ncoarse-grained texture\nsolid\nnot made by organisms',
+% 'task': 'closed choice',
+% 'grade': 'grade8',
+% 'subject': 'natural science',
+% 'topic': 'earth-science',
+% 'category': 'Rocks and minerals',
+% 'skill': 'Identify rocks and minerals',
+% 'lecture': 'Minerals are the building blocks of rocks.
 A rock can be made of one or more minerals.\nMinerals and rocks have the following properties:\nProperty | Mineral | Rock\nIt is a solid.
@@
 So, the arrangement of atoms or molecules in different pieces of the same type of rock may be different!',
- 'solution': 'The properties of diorite match the properties of a rock. So, diorite is a rock.'}
-\end{verbatim}
+% 'solution': 'The properties of diorite match the properties of a rock. So, diorite is a rock.'}
+%\end{verbatim}
 
@@
 
-For other datapoints, however, the image data can be missing, although required to make an answer, e.g.
-\begin{verbatim}
-{'image': None,
- 'question': 'What information supports the conclusion that Rick inherited this trait?',
- 'choices': ["Rick's coworker also has curly hair.",
-  "Rick's biological father has curly hair.",
-  'Rick and his biological parents have brown hair.'],
- 'answer': 1,
- 'hint': 'Read the description of a trait.\nRick has curly hair.',
- 'task': 'closed choice',
- 'grade': 'grade7',
- 'subject': 'natural science',
- 'topic': 'biology',
- 'category': 'Genes to traits',
- 'skill': 'Inherited and acquired traits: use evidence to support a statement',
- 'lecture': "Organisms, including people, have both inherited and acquired traits.
+For other datapoints, however, the image data can be missing, although required to make an answer.
+%, e.g.
+%\begin{verbatim}
+%{'image': None,
+% 'question': 'What information supports the conclusion that Rick inherited this trait?',
+% 'choices': ["Rick's coworker also has curly hair.",
+%  "Rick's biological father has curly hair.",
+%  'Rick and his biological parents have brown hair.'],
+% 'answer': 1,
+% 'hint': 'Read the description of a trait.\nRick has curly hair.',
+% 'task': 'closed choice',
+% 'grade': 'grade7',
+% 'subject': 'natural science',
+% 'topic': 'biology',
+% 'category': 'Genes to traits',
+% 'skill': 'Inherited and acquired traits: use evidence to support a statement',
+% 'lecture': "Organisms, including people, have both inherited and acquired traits.
 Inherited and acquired traits are gained in different ways.\nInherited traits are passed down from biological parents to their offspring through genes.
@@
 Parents do not pass acquired traits down to their offspring.",
- 'solution': ''}
-\end{verbatim}
+% 'solution': ''}
+%\end{verbatim}
 
@@
 
-\paragraph{Metrics}
-We'll evaluate model's performance with question answering accuracy domain-wise in order to have a fair comparison with leaderboards.
+%\paragraph{Metrics}
+%We'll evaluate model's performance with question answering accuracy domain-wise in order to have a fair comparison with leaderboards.
 The reasoning steps can be evaluated with semantic similarity metrics (e.g.
 adopted from machine translation) such as BLEU, METEOR, ROUGE, BLANC, and BERTScore and more.
+%\subsection{Multiple choice Evaluation}
+%Owing to the simplicity of the test format, only accuracy score is computed, following original evaluation strategy by \cite{lu2022learn}. This metric is considered of the highest importantce 
+%\subsection{Answer Reasoning Evaluation}
+%Due to the pecularity of scientific texts and approaches to the evaluation of automatically generated texts, the following evaluation approaches were chosen.
+Our validation process is consisted of the metrics introduced in the original paper \cite{lu2022learn}  with a few additional metrics and techniques that we consider of high importance in regards to the scientific textual data.
 %, but the main metric we'll rely on is accuracy for question answering.
+%removing this as it will be replaced by better text from Dasha
 
 \section{Experiments}
+\label{experiments}
 
-\subsection{Zero-shot and Few-shot Benchmarking}
-We start by zero-shot benchmarking existing multimodal models in three simple prompt settings, partly repeating with the official evaluation suite:
+\subsection{Setup}
+We start by zero-shot benchmarking existing multimodal models in four simple prompt settings, partly repeating with the official evaluation suite:
 \begin{enumerate}
-	\item \textbf{question - choices - hint - image - task}, with the following prompt:
-	\begin{verbatim}{f'Question: {question}\n Task: {task}\n Choices: {choices}\n Hint: {hint}'}\end{verbatim}
-	\item \textbf{question - choices - hint - image - task + lecture}, with the following prompt:
-	\begin{verbatim}{f'Question: {question}\n Task: {task}\n Choices: {choices}\n Hint: {hint}\n 
-	Lecture: {lecture}'}\end{verbatim}
-	\item \textbf{question - choices - hint - image - task + lecture + solution}, with the following prompt:
-	\begin{verbatim}{f'Question: {question}\n Task: {task}\n Choices: {choices}\n Hint: {hint}\n 
-	Lecture: {lecture}\n Solution: {solution}'}\end{verbatim}
+	\item \textbf{question - choices - hint - image - task}
+	\item \textbf{question - choices - hint - image - task + lecture}
+	\item \textbf{question - choices - hint - image - task + lecture + solution}
+	\item \textbf{question - choices - hint - image - task + solution}
 \end{enumerate}
@@
 
-\section{Benchmark}
+\subsection{Metrics}
 % description of metrics, why they were used + analytics of results
-\subsection{Accuracy}
+\subsubsection{Accuracy}
 The highest performance was observed in settings that included solutions, indicating that the models were generally capable of extracting relevant information effectively.
@@
 
-\subsection{Cosine similarity}
-\subsection{BLEU-1}
+\subsubsection{Cosine similarity}
+\subsubsection{BLEU-1}
 Across settings, higher BLEU-1 scores were observed in settings 3 and 4 (those incorporating solutions and/or lecture information), indicating that models performed better with richer contextual data.
@@
 
-\subsection{BLEU-4}
+\subsubsection{BLEU-4}
 Scores were consistently higher in settings 3 and 4, particularly for  models, demonstrating their ability to produce more contextually precise outputs when both lecture and solution information were provided.
@@
 
-\subsection{ROUGE}
+\subsubsection{METEOR}
+TODO
+
+\subsubsection{ROUGE}
 The highest ROUGE scores were observed in setting 3, where both lectures and solutions were available, highlighting the importance of comprehensive input for producing informative responses.
@@
 
-\subsection{Cosine similarity}
+\subsubsection{Cosine similarity}
 Semantic alignment with the reference is assessed by the similarity metric.
@@
 
-\subsection{Summary}
+\subsubsection{Summary}
 Overall, the findings demonstrate how important enriched contextual data, such as lectures and solutions, in enhancing model performance across all metrics.
@@
 
+\subsection{Zero-shot Benchmarking on Validation}
 \subsubsection{Local Models from HuggingFace}
@@
 
-\paragraph{MiniGPT-4} This is a small multimodal model designed from ?? \cite{zhu2023minigpt}. It offers to build the model based on Llama or Vicuna. 
-\textbf{Unsuccessful experiment}: We tried to resort to an unofficial implementation: \texttt{wangrongsheng/MiniGPT-4-LLaMA-7B}. Loading the model was not successful (no preprocessor\_config.json). 
-Another issue: using BlipForConditionalGeneration and BlipProcessor: "You are using a model of type llama to instantiate a model of type blip.
+%Unsuccessful attempts can be mentioned in appendix I guess
+%\paragraph{MiniGPT-4} This is a small multimodal model designed from ?? \cite{zhu2023minigpt}. It offers to build the model based on Llama or Vicuna. 
+%\textbf{Unsuccessful experiment}: We tried to resort to an unofficial implementation: \texttt{wangrongsheng/MiniGPT-4-LLaMA-7B}. Loading the model was not successful (no preprocessor\_config.json). 
+%Another issue: using BlipForConditionalGeneration and BlipProcessor: "You are using a model of type llama to instantiate a model of type blip.
 This is not supported for all configurations of models and can yield errors.".
@@
 
-\paragraph{BLIP-2}
+%\paragraph{BLIP-2}
 
-\paragraph{LLaMA models}
+%\paragraph{LLaMA models} -- forbidden in EU, ignore
 
-\paragraph{Gemma2}
+%\paragraph{Gemma2} -- Paligemma has seen ScienceQA
 
-\paragraph{Qwen models} e.g.  Qwen2.5-Coder
+\paragraph{Qwen models} 
+%e.g.  Qwen2.5-Coder
+Current experiment
 
-\paragraph{}
+\paragraph{Pixtral}
 
@@
 
-The metrics used for evaluation at this point are: accuracy, \texttt{BLEU-1}, \texttt{BLEU-4}, \texttt{ROUGE}, similarity.
+%The metrics used for evaluation at this point are: accuracy, \texttt{BLEU-1}, \texttt{BLEU-4}, \texttt{ROUGE}, similarity.
 
-\section{Approximate Timeline}
-\begin{enumerate}
-	\item \textbf{Now - mid November}: 
-		\begin{itemize}
-  			\item Preliminary baseline tests: compare "pure LLM"s relying only on commonsense reasoning / only on pretraining (zero-shot, few-shot).
-			\item Stronger baseline – same LLM with better prompting techniques like CoT/Self-Ask.
-		\end{itemize}
-	\item \textbf{mid November - mid December}:
-		\begin{itemize}
-  			\item Construct agentic / RAG pipeline with “big smart” model + additional resources.
-  			\item Optional: soft-prompting for better workflow to distill from.
-		\end{itemize}
-	\item \textbf{mid December - mid January}:
-  		\begin{itemize}
-			\item Knowledge distillation experiments.
-		\end{itemize}
-\end{enumerate}
+%\section{Approximate Timeline}
+%\begin{enumerate}
+%	\item \textbf{Now - mid November}: 
+%		\begin{itemize}
+%  			\item Preliminary baseline tests: compare "pure LLM"s relying only on commonsense reasoning / only on pretraining (zero-shot, few-shot).
+%			\item Stronger baseline – same LLM with better prompting techniques like CoT/Self-Ask.
+%		\end{itemize}
+%	\item \textbf{mid November - mid December}:
+%		\begin{itemize}
+ % 			\item Construct agentic / RAG pipeline with “big smart” model + additional resources.
+%  			\item Optional: soft-prompting for better workflow to distill from.
+%		\end{itemize}
+%	\item \textbf{mid December - mid January}:
+%  		\begin{itemize}
+%			\item Knowledge distillation experiments.
+%		\end{itemize}
+%\end{enumerate}
 
-\printbibliography
+\section{Knowledge Distillation \& Soft Prompting}
+\label{knowledge-distillation}
+
+\section{Error Analysis}
+\label{error-analysis}
+
+\section{Conclusion}
+\label{Conclusion}
+
+%\printbibliography
+\bibliography{example_paper}
+\bibliographystyle{icml2021}
 

### 2024-12-09 docs/paper/main.tex

@@
 
-%\paragraph{Metrics}
-%We'll evaluate model's performance with question answering accuracy domain-wise in order to have a fair comparison with leaderboards.
+\paragraph{Metrics}
+We evaluate model's performance with question answering accuracy domain-wise in order to have a fair comparison with leaderboards.
 The reasoning steps can be evaluated with semantic similarity metrics (e.g.
-adopted from machine translation) such as BLEU, METEOR, ROUGE, BLANC, and BERTScore and more.
-%\subsection{Multiple choice Evaluation}
-%Owing to the simplicity of the test format, only accuracy score is computed, following original evaluation strategy by \cite{lu2022learn}. This metric is considered of the highest importantce 
-%\subsection{Answer Reasoning Evaluation}
-%Due to the pecularity of scientific texts and approaches to the evaluation of automatically generated texts, the following evaluation approaches were chosen.
+adopted from machine translation) such as BLEU, METEOR, ROUGE, and cosine similarity and more.
+\subsection{Multiple choice Evaluation}
+Owing to the simplicity of the test format, only accuracy score is computed, following original evaluation strategy by \cite{lu2022learn}.
+It was determined that the computation of the overall score should assign the highest weight (\textbf{50\%}) to the ability to provide correct answers (accuracy score), as this represents the most critical aspect of performance evaluation.
+\subsection{Answer Reasoning Evaluation}
+Due to the pecularity of scientific texts and approaches to the evaluation of automatically generated texts, the following evaluation approaches were chosen.
 Our validation process is consisted of the metrics introduced in the original paper \cite{lu2022learn}  with a few additional metrics and techniques that we consider of high importance in regards to the scientific textual data.
-%, but the main metric we'll rely on is accuracy for question answering.
-%removing this as it will be replaced by better text from Dasha
+
+\subsubsection{BLEU}
+BLEU-score (\cite{papineni2002bleu}) can measure how closely the model's generated explanation aligns with the human-authored explanation.
+BLEU-1 measures if the model uses the right scientific terms or key words (e.g., "photosynthesis," "temperature").
+However, it ignores word order and logical progression, so it can’t evaluate reasoning or explanation quality.
+That is why BLEU-4 score is also computed in order to capture both vocabulary and the arrangement of words into meaningful phrases, evaluating fluency and coherence.
+Scientific explanations can often convey the same idea with different wording, making semantic similarity hard to capture.
+BLEU might penalize valid but rephrased or concise answers and it does not account for deeper meaning or logical correctness, which are crucial in reasoning tasks.
+Since this metric can give valuable insights on the generated outputs, we do not consider these scores as the most important and both BLEU-1 and BLEU-4 scores have \textbf{5\%} impact.
+
+\subsubsection{ROUGE}
+ROUGE-L score, following evaluation strategy in \cite{lu2022learn}, helps analyze the content recall and sequence preservation of model-generated explanations.
+They highlight whether a model captures key scientific concepts and produces explanations that align with reference reasoning.
+However, ROUGE should be used in conjunction with other metrics to fully evaluate the logical and factual quality of generated reasoning.
+ROUGE-L will highlight the mismatch in logical flow, even though some content overlaps in reference and candidate sentences (\ref{example:1}, \ref{example:2}).
+It lead to the conclusion that this metric would not drastically affect the overall score and contribute (\textbf{5\%}).
+\begin{example}
+\label{example:1} 
+\begin{itemize}
+    \item \textbf{Reference:} Water evaporates, cools, and condenses to form clouds.
+    \item \textbf{Candidate:} Water cools, evaporates, and forms clouds.
+\end{itemize}
+\end{example}
+\begin{example}
+\label{example:2}
+\begin{itemize}
+    \item \textbf{Reference:} Plants absorb sunlight to produce energy.
+    \item \textbf{Candidate:} Sunlight is absorbed by plants to create energy.
+\end{itemize}
+\end{example}
+\subsubsubsection{Cosine similarity with Sentence Transformers}
+Cosine similarity using SentenceTransformer embeddings provides a robust method for evaluating model outputs in the ScienceQA dataset, focusing on the semantic alignment of generated explanations with references.
+It captures nuanced relationships and conceptual understanding, which gives more understanding of the generated outputs in comparison to ROUGE and BLEU scores and that is why it has more importance in final score computation.
+However, due to domain specificity it can miss some intricate cases like still considering high similarity of two sentences, even though it the reference can contain crucial information about carbon dioxide, while in the candidate sentence the answer was given in relation to oxygen reproduction.
+\subsubsection{METEOR}
+METEOR (\cite{banerjee2005meteor}) evaluates text similarity based on precision, recall, and alignment of semantically meaningful components like synonyms and stems, making it well-suited for capturing the nuanced language used in scientific reasoning.
+By incorporating a weighted F-score and penalties for excessive mismatches, METEOR provides a balanced evaluation of fluency and accuracy, crucial for assessing LLM-generated answers on the Science QA Dataset.
+Due to these balanced text evaluation, this metric has \textbf{10\%} impact on the overall score.
+% \subsubsection{Cosine similarity with SciBERT}
+% We extended the evaluation of cosine similarity using SciBERT (\cite{beltagy2019scibert}) embeddings.
+These vectors are more consistent with the domain-specific terminology and reasoning patterns present in the Science QA Dataset because it has been properly pre-trained on scientific literature.
+% SciBERT generates more accurate similarity embeddings because it better reflects the complex relationships between scientific concepts than general-purpose models like those in Sentence Transformers.
+% When assessing the logical and factual coherence of scientific reasoning, the structural and contextual semantics specific to scientific texts are reflected in the embeddings produced by SciBERT.
+% Assessments for jobs requiring sophisticated scientific understanding are more reliable when SciBERT is used for cosine similarity since it guarantees that the evaluation metric is domain-specific.
+
+\subsubsection{Cosine similarity with Sentence Transformers}
+Sentence Transformer models are optimized for sentence-level embeddings, which capture the overall contextual similarity of answers more effectively, making them suitable for general-purpose evaluation tasks like scientific reasoning.
+Unlike SciBERT, which is tailored to scientific text and could have been used as sentence embedder, Sentence Transformers are pre-trained on diverse datasets, enabling them to generalize better to variations in phrasing and style present in the Science QA Dataset.
+Their ability to handle broader linguistic nuances ensures that the evaluation accounts for both semantic correctness and the logical flow of responses, which is essential when assessing reasoning across a wide range of scientific topics.
+That is the reason for using Sentence Transformer model for measurements of cosine similarity between the lecture material and annotated explanations and models' outputs with impact weight of \textbf{15\%}.
 

### 2024-12-10 docs/paper/main.tex
Dataset section

@@
 This study is based on the ScienceQA dataset \cite{lu2022learn}.
-The dataset includes a variety of science-related multimodal multiple-choice questions together with annotations of the answers that provide relevant lectures and explanations in Nature Science, Language Science and Social Science.
-
-%Data points are structured as follows:
-%\begin{verbatim}
-%{'image': <PIL.PngImagePlugin.PngImageFile image mode=RGB size=200x202>,
-% 'question': 'Is diorite a mineral or a rock?',
-% 'choices': ['rock', 'mineral'],
-% 'answer': 0,
-% 'hint': 'Diorite has the following properties:\nno fixed crystal structure\nnaturally occurring\nnot a pure substance\ncoarse-grained texture\nsolid\nnot made by organisms',
-% 'task': 'closed choice',
-% 'grade': 'grade8',
-% 'subject': 'natural science',
-% 'topic': 'earth-science',
-% 'category': 'Rocks and minerals',
-% 'skill': 'Identify rocks and minerals',
-% 'lecture': 'Minerals are the building blocks of rocks.
-A rock can be made of one or more minerals.\nMinerals and rocks have the following properties:\nProperty | Mineral | Rock\nIt is a solid.
-| Yes | Yes\nIt is formed in nature.
-| Yes | Yes\nIt is not made by organisms.
-| Yes | Yes\nIt is a pure substance.
-| Yes | No\nIt has a fixed crystal structure.
-| Yes | No\nYou can use these properties to tell whether a substance is a mineral, a rock, or neither.\nLook closely at the last three properties:\nMinerals and rocks are not made by organisms.\nOrganisms make their own body parts.
-For example, snails and clams make their shells.
-Because they are made by organisms, body parts cannot be  minerals or rocks.\nHumans are organisms too.
-So, substances that humans make by hand or in factories are not minerals or rocks.\nA mineral is a pure substance, but a rock is not.\nA pure substance is made of only one type of matter.
-Minerals are pure substances, but rocks are not.
-Instead, all rocks are mixtures.\nA mineral has a fixed crystal structure, but a rock does not.\nThe crystal structure of a substance tells you how the atoms or molecules in the substance are arranged.
-Different types of minerals have different crystal structures, but all minerals have a fixed crystal structure.
-This means that the atoms and molecules in different pieces of the same type of mineral are always arranged the same way.\nHowever, rocks do not have a fixed crystal structure.
-So, the arrangement of atoms or molecules in different pieces of the same type of rock may be different!',
-% 'solution': 'The properties of diorite match the properties of a rock. So, diorite is a rock.'}
-%\end{verbatim}
-
-For some datapoints, the image data is missing, while the answer can be deduced from text alone (e.g.
-questions on 'subject': 'language science', 'topic': 'figurative-language', 'skill': 'Interpret figures of speech' (e.g.
-irony recognition).
-
-For other datapoints, however, the image data can be missing, although required to make an answer.
-%, e.g.
-%\begin{verbatim}
-%{'image': None,
-% 'question': 'What information supports the conclusion that Rick inherited this trait?',
-% 'choices': ["Rick's coworker also has curly hair.",
-%  "Rick's biological father has curly hair.",
-%  'Rick and his biological parents have brown hair.'],
-% 'answer': 1,
-% 'hint': 'Read the description of a trait.\nRick has curly hair.',
-% 'task': 'closed choice',
-% 'grade': 'grade7',
-% 'subject': 'natural science',
-% 'topic': 'biology',
-% 'category': 'Genes to traits',
-% 'skill': 'Inherited and acquired traits: use evidence to support a statement',
-% 'lecture': "Organisms, including people, have both inherited and acquired traits.
-Inherited and acquired traits are gained in different ways.\nInherited traits are passed down from biological parents to their offspring through genes.
-Genes are pieces of hereditary material that contain the instructions that affect inherited traits.
-Offspring receive their genes, and therefore gain their inherited traits, from their biological parents.
-Inherited traits do not need to be learned.\nAcquired traits are gained during a person's life.
-Some acquired traits, such as riding a bicycle, are gained by learning.
-Other acquired traits, such as scars, are caused by the environment.
-Parents do not pass acquired traits down to their offspring.",
-% 'solution': ''}
-%\end{verbatim}
-
-Apart from the image field, missing values can occur in lecture and solution fields. The statistics in table \ref{table:dataset-stats} give an overview on the amount of problematic datapoints.
-
-Finally, hint can be missing, e.g.
-the text-only task 367 on subject language science: "What is the source of allusion in the sentence below?
-Edward picked up his pace on the trail as his spidey sense began to tingle" with choices ["Italian history", "a comic book"], or the text-only task 3 on the topic of capitalization: "Which correctly shows the title of a play?" with choices ["A breath of Fresh Air", "A Breath of Fresh Air"].
-
-\begin{table}[h!]
-\centering
-\begin{tabular}{ll}\toprule
-	missing & frequency \\\midrule
-	solution & 9.3\% \\
-	lecture & 15.7\% \\
-	solution and/or lecture & 24.2\% \\
-	image & 50.6\%\\\bottomrule
-\end{tabular}
-\caption{Defect datapoints in dataset}
-\label{table:dataset-stats}
-\end{table}
-
-\paragraph{Retained data} As 50.6\% of the datapoints are missing an image, we decide to process both text-only and text+image datapoints similarly with multimodal models, attaching the image to the prompt if it is available.
-For now, we will only work with datapoints with non-empty solution AND with non-empty lecture. All the other datapoints are disregarded,
-%\begin{verbatim}
-%filtered\_validation\_data = validation\_data.filter(lambda example: example["lecture"] != "" and example["solution"] != "")
-%\end{verbatim}
-which results in \textbf{3216} datapoints kept in validation data.
+The ScienceQA dataset is a benchmark for multimodal reasoning in science, consisting of over 21,000 questions across topics like natural science, social science and language science.
+Each question includes textual prompts, optionally visual aids (e.g., diagrams, charts), answer options, detailed explanations and lecture context.
+The dataset spans various difficulty levels, enabling evaluation of models on both basic and advanced scientific reasoning, reaching from elementary to high school level questions.
+ScienceQA supports tasks such as multimodal comprehension, answer explanation generation, and educational AI development.
+Its high-quality annotations and diverse content make it an ideal benchmark for assessing models in science education and reasoning.
+In our experiments, we leveraged the dataset to assess knowledge distillation under diverse scientific scenarios, focusing on the ability to distill the reasoning capabilities.
+Irregularities:
+For some datapoints, the image data is missing, while the answer can be deduced from text alone.
+Apart from the image field, missing values can occur in lecture and solution fields.
+As about 50\% of the datapoints are missing an image, we decide to process both text-only and text+image datapoints similarly with multimodal models, attaching the image to the prompt if it is available.
 

### 2024-12-10 docs/paper/main.tex
text_and_data

@@
 Owing to the simplicity of the test format, only accuracy score is computed, following original evaluation strategy by \cite{lu2022learn}.
-It was determined that the computation of the overall score should assign the highest weight (\textbf{50\%}) to the ability to provide correct answers (accuracy score), as this represents the most critical aspect of performance evaluation.
+It was determined that the computation of the overall score should assign the highest weight (\textbf{60\%}) to the ability to provide correct answers (accuracy score), as this represents the most critical aspect of performance evaluation.
 \subsection{Answer Reasoning Evaluation}
@@
 By incorporating a weighted F-score and penalties for excessive mismatches, METEOR provides a balanced evaluation of fluency and accuracy, crucial for assessing LLM-generated answers on the Science QA Dataset.
-Due to these balanced text evaluation, this metric has \textbf{10\%} impact on the overall score.
+Due to these balanced text evaluation, this metric has \textbf{15\%} impact on the overall score.
 % \subsubsection{Cosine similarity with SciBERT}
@@
 Their ability to handle broader linguistic nuances ensures that the evaluation accounts for both semantic correctness and the logical flow of responses, which is essential when assessing reasoning across a wide range of scientific topics.
-That is the reason for using Sentence Transformer model for measurements of cosine similarity between the lecture material and annotated explanations and models' outputs with impact weight of \textbf{15\%}.
+That is the reason for using Sentence Transformer model for measurements of cosine similarity between the lecture material and annotated explanations and models' outputs with impact weight of \textbf{10\%}.
 
@@
 Among these, settings combining both lecture and solution information typically yielded slightly better results than those with solutions alone.
-`GPT-4` consistently received highest ratings, especially in scenarios in which lecture and/or solution information was included.
-The Gemini family of models showed better robustness in many scenarios and came in second overall.
+`GPT-4` and `Gemini-1.5-flash-8B` consistently received highest ratings, especially in scenarios in which lecture and/or solution information was included.
+The Gemini family of models and  `Pixtral` model showed better robustness in many scenarios and came in second overall.
 Notably, Gemini models achieved an average accuracy advantage of 5\% over GPT-family models in the "pure" task context (without lecture or solution information).
@@
 \subsubsection{Cosine similarity}
+Cosine similarity highlights the semantic alignment between the model outputs and the correct answers.
+`Gemini-1.5` consistently achieves the highest similarity scores, particularly in settings 3 and 4, indicating strong alignment with the reference answers.
+`GPT-4` variants maintain steady scores across all settings, demonstrating robustness, while `Pixtral` and `LLaVAModel1.5-7b` show more variability, suggesting sensitivity to specific settings.
+Overall, cosine similarity shows that `Gemini-1.5` excels in semantic understanding across the dataset.
 \subsubsection{BLEU-1}
 Across settings, higher BLEU-1 scores were observed in settings 3 and 4 (those incorporating solutions and/or lecture information), indicating that models performed better with richer contextual data.
-The `GPT-4o-mini` model demonstrated superior performance, surpassing `GPT-4`, in achieving the highest results.
+The `Llave-1.5-flash-8B` model demonstrated superior performance, surpassing `GPT-4`, in achieving the highest results.
 Lower scores in settings 1 and 2 suggest that models, particularly `LLaVAModel1.5-7b`, may have had trouble correctly aligning their outputs with reference words when given less auxiliary information, as indicated by lower scores in settings 1 and 2.
@@
 \subsubsection{BLEU-4}
-Scores were consistently higher in settings 3 and 4, particularly for  models, demonstrating their ability to produce more contextually precise outputs when both lecture and solution information were provided.
 Gemini models had higher scores in comparison to other models, which represents a better capability of information extraction.
-`LlaVAModel`, at the same time, gained the highest score in the setting without solutions, which indicates the ability of extracting knowledge from massive textual data.
+`GPT-4o-mini`, at the same time, gained the highest score in the setting without solutions, which indicates the ability of extracting knowledge from massive textual data.
 
 \subsubsection{METEOR}
-TODO
+METEOR scores reveal how well the models capture fluency, grammar, and word-level alignment with the references.
+Similar to cosine similarity, `Gemini-1.5` leads with the highest METEOR scores.
+Notably, `Pixtral` achieves relatively high METEOR scores in settings 3 and 4, despite moderate cosine similarity, highlighting its ability to produce fluent outputs.
+When concrete tasks without any additional helpful material are given to the models as an input, `gpt-4o-mini` model shows a capability of generating concise, fluent answers with correct relevant terminology and explanations.
 
@@
 
-\subsubsection{Cosine similarity}
-Semantic alignment with the reference is assessed by the similarity metric.
-Gemini and its variants outperformed other models, demonstrating their great capacity to capture the semantic structure of enriched input.
-Similarity scores were highest in settings 3 and 4.
-LLaVAModel1.5-7b, on the other hand, received the lowest score in setting 1, indicating that it mostly depends on lecture or solution data to generate outputs that are semantically accurate.
-This result emphasises how more data affects the resulting content's semantic integrity.
-`GPT-4o-mini` and `LLaVAModel` performed better than other models in environments lacking solution information, indicating that they can produce reliable results even with limited  contextual guidance.
+\subsubsection{Overall score}
+The overall results indicate that the `GPT` and `Gemini` model families demonstrate exceptional ability in extracting accurate answers.
+Notably, the `Gemini` models exhibited superior performance on datasets lacking relevant lecture information.
+The second-best performance on such "pure" datasets was achieved by `GPT-4`, with only a slight difference in overall score.
 
-\subsubsection{Summary}
-Overall, the findings demonstrate how important enriched contextual data, such as lectures and solutions, in enhancing model performance across all metrics.
-The GPT-4o-mini model outperformed GPT-4 in multiple instances, indicating the efficacy of smaller, more targeted designs, even if GPT-4 continuously received excellent accuracy and informativeness scores.
-While LLaVAModel1.5-7b failed without extra input, highlighting its dependence on extensive contextual knowledge to function well, the Gemini family of models demonstrated remarkable robustness and semantic alignment, particularly in enriched situations.
-Lecture information doesn’t really affect models’ performance, it can even decrease the ratio of correct answers.
+\begin{figure}
+    \centering
+\includegraphics[width=0.8\textwidth]{images/overall_metric.pdf} 
+    \caption{Overals l metrics of various LLMs.}
+    \label{fig:metrics}
+\end{figure}
+
+\subsection{Summary}
+The benchmarking analysis highlights the superior performance of the `Gemini` and `GPT` model families, particularly in scenarios with combined lecture and solution data.
+The overall score indicates that both `Gemini` and `GPT` models excel at extracting accurate answers, with `Gemini` models showing a notable advantage on datasets lacking lecture information.
+`GPT-4` ranked second in performance on these "pure" datasets, with only a slight difference in overall score.
+`Pixtral `demonstrated strong fluency despite moderate semantic alignment, while `LLaVAModel1.5-7b` showed the lowest accuracy, particularly in minimal-context settings.
 

### 2024-12-28 prefix_tuning.py
Prefix Tuning

@@ -36,7 +36,5 @@ class PrefixTuningModel(nn.Module):
         # Modify attention mask for prefix
         prefix_mask = torch.ones((inputs["input_ids"].size(0), self.prefix_tuning.prefix_length), device=inputs["input_ids"].device)
         attention_mask = torch.cat([prefix_mask, inputs["attention_mask"]], dim=1)
-        print(f"Input Embed dims: {inputs_embeds.size()}")
-        print(f"Label: {labels.size()}")
 
         return self.model(inputs_embeds=inputs_embeds, attention_mask=attention_mask, pixel_values=inputs["pixel_values"], labels=labels)

### 2024-12-28 bwcluster/inference_model_QTCH.sh
bwcluster vllm inference

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --job-name=qtchPixtralSciQA
+#SBATCH --output=job_output_qtch_%j.log
+#SBATCH --gres=gpu:4
+#SBATCH --time=00:30:00
+#SBATCH --partition=dev_gpu_4_a100
+export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
+source ~/fmenv/bin/activate
+python inference_qwen_model.py --ngpus 4 --setting QTCH --resume 0

### 2024-12-28 bwcluster/inference_model_QTCHL.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --job-name=qtchlPixtralSciQA
+#SBATCH --output=job_output_qtchl_%j.log
+#SBATCH --gres=gpu:4
+#SBATCH --time=00:30:00
+#SBATCH --partition=dev_gpu_4_a100
+export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
+source ~/fmenv/bin/activate
+python inference_qwen_model.py --ngpus 4 --setting QTCHL --resume 0

### 2024-12-28 bwcluster/inference_model_QTCHLS.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --job-name=qtchlsPixtralSciQA
+#SBATCH --output=job_output_qtchls_%j.log
+#SBATCH --gres=gpu:4
+#SBATCH --time=00:50:00
+#SBATCH --partition=gpu_4
+export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
+source ~/fmenv/bin/activate
+python inference_model.py --ngpus 4 --setting QTCHLS --resume 208

### 2024-12-28 bwcluster/inference_model_QTCHS.sh

@@ -1,0 +1,10 @@
+#!/bin/bash
+#SBATCH --job-name=PixtralSciQA
+#SBATCH --output=job_output_%j.log
+#SBATCH --gres=gpu:4
+#SBATCH --time=01:00:00
+#SBATCH --partition=gpu_4
+export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
+source ~/fmenv/bin/activate
+python inference_model.py --ngpus 4 --setting QTCH --resume 320
+python inference_model.py --ngpus 4 --setting QTCHLS --resume 207

### 2024-12-28 bwcluster/inference_qwen_model.py

@@ -1,0 +1,259 @@
+# Script to download and inference with mistral models
+
+import csv
+import os
+import pandas as pd
+from collections import defaultdict
+from transformers import Qwen2VLForConditionalGeneration, AutoTokenizer, AutoProcessor
+from qwen_vl_utils import process_vision_info
+
+import argparse
+import logging
+from datetime import datetime
+
+# Generate timestamped filename
+start_time = datetime.now().strftime("%Y%m%d_%H%M%S")
+log_filename = f"inference_{start_time}.log"
+
+# Configure logging
+logging.basicConfig(
+    level=logging.INFO,
+    format="%(asctime)s - %(levelname)s - %(message)s",
+    handlers=[
+        logging.FileHandler(log_filename),  # Log to file
+        logging.StreamHandler()            # Also log to console (optional)
+    ]
+)
+
+from vllm import LLM, TokensPrompt
+from vllm import EngineArgs, LLMEngine
+from vllm.sampling_params import SamplingParams
+from vllm.multimodal import MultiModalDataBuiltins
+
+from huggingface_hub import login
+from datasets import load_dataset
+
+## UTILS: LOAD MODEL
+# login to hf -- mistral does not allow unauthorized downloads
+def login_to_hf():
+    try:
+        with open("hf_token", "r") as f:
+            token = f.read().strip()
+    except FileNotFoundError:
+        raise FileNotFoundError("HuggingFace token file not found. Please ensure 'hf_token' exists.")
+        return
+
+    login(token=token)
+
+    logging.info("Login with token: done")
+
+# download or load from cache the model, put on GPU or distribute among GPUs, prepare CUDA graph
+def prepare_model(model_name, tensor_parallel_size, cpu_offload_gb, max_model_len):
+    # from documentation - default: Load the model on the available device(s)
+    # model = Qwen2VLForConditionalGeneration.from_pretrained(
+    #     "Qwen/Qwen2-VL-2B-Instruct", torch_dtype="auto", device_map="auto"
+    # )
+    llm = LLM(model=model_name, 
+              tokenizer_mode="auto",
+              task="generate",
+              tensor_parallel_size=tensor_parallel_size,  # Can be adjusted for multi-GPU setup
+              cpu_offload_gb=cpu_offload_gb,
+              dtype="float16",
+              # dtype="bfloat16", #"float32",
+              max_model_len=max_model_len # Setting to same number as SamplingParams
+             )
+
+    logging.info("Load model: done")
+    
+    return llm
+
[…]

### 2024-12-31 bwcluster/inference_model_QTCH.sh
qwen2 zero-shot scoring

@@ -1,9 +1,9 @@
 #!/bin/bash
-#SBATCH --job-name=qtchPixtralSciQA
+#SBATCH --job-name=qtchSciQA
 #SBATCH --output=job_output_qtch_%j.log
-#SBATCH --gres=gpu:4
-#SBATCH --time=00:30:00
-#SBATCH --partition=dev_gpu_4_a100
+#SBATCH --gres=gpu:1
+#SBATCH --time=00:20:00
+#SBATCH --partition=gpu_4
 export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
 source ~/fmenv/bin/activate
-python inference_qwen_model.py --ngpus 4 --setting QTCH --resume 0
+python inference_qwen_model.py --ngpus 1 --setting QTCH --resume 3161

### 2024-12-31 bwcluster/inference_model_QTCHL.sh

@@ -1,9 +1,9 @@
 #!/bin/bash
-#SBATCH --job-name=qtchlPixtralSciQA
+#SBATCH --job-name=qtchlSciQA
 #SBATCH --output=job_output_qtchl_%j.log
-#SBATCH --gres=gpu:4
+#SBATCH --gres=gpu:1
 #SBATCH --time=00:30:00
-#SBATCH --partition=dev_gpu_4_a100
+#SBATCH --partition=gpu_4
 export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
 source ~/fmenv/bin/activate
-python inference_qwen_model.py --ngpus 4 --setting QTCHL --resume 0
+python inference_qwen_model.py --ngpus 1 --setting QTCHL --resume 3054

### 2024-12-31 bwcluster/inference_model_QTCHLS.sh

@@ -1,9 +1,9 @@
 #!/bin/bash
-#SBATCH --job-name=qtchlsPixtralSciQA
+#SBATCH --job-name=qtchlsSciQA
 #SBATCH --output=job_output_qtchls_%j.log
-#SBATCH --gres=gpu:4
-#SBATCH --time=00:50:00
+#SBATCH --gres=gpu:1
+#SBATCH --time=00:30:00
 #SBATCH --partition=gpu_4
 export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
 source ~/fmenv/bin/activate
-python inference_model.py --ngpus 4 --setting QTCHLS --resume 208
+python inference_qwen_model.py --ngpus 1 --setting QTCHLS --resume 3019

### 2024-12-31 bwcluster/inference_model_QTCHS.sh

@@ -1,10 +1,9 @@
 #!/bin/bash
-#SBATCH --job-name=PixtralSciQA
-#SBATCH --output=job_output_%j.log
-#SBATCH --gres=gpu:4
-#SBATCH --time=01:00:00
+#SBATCH --job-name=qtchsSciQA
+#SBATCH --output=job_output_qtchs_%j.log
+#SBATCH --gres=gpu:1
+#SBATCH --time=00:20:00
 #SBATCH --partition=gpu_4
 export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
 source ~/fmenv/bin/activate
-python inference_model.py --ngpus 4 --setting QTCH --resume 320
-python inference_model.py --ngpus 4 --setting QTCHLS --resume 207
+python inference_qwen_model.py --ngpus 1 --setting QTCHS --resume 3190

### 2024-12-31 bwcluster/inference_qwen_model.py

@@ -33,6 +33,8 @@ from vllm.multimodal import MultiModalDataBuiltins
 from huggingface_hub import login
 from datasets import load_dataset
 
+from PIL import Image
+
 ## UTILS: LOAD MODEL
 # login to hf -- mistral does not allow unauthorized downloads
 def login_to_hf():
@@ -59,8 +61,7 @@ def prepare_model(model_name, tensor_parallel_size, cpu_offload_gb, max_model_le
               tensor_parallel_size=tensor_parallel_size,  # Can be adjusted for multi-GPU setup
               cpu_offload_gb=cpu_offload_gb,
               dtype="float16",
-              # dtype="bfloat16", #"float32",
-              max_model_len=max_model_len # Setting to same number as SamplingParams
+              # max_model_len=max_model_len 
              )
 
     logging.info("Load model: done")
@@ -123,51 +124,41 @@ def process_datapoint_at_setting(example, idx, setting, llm, model_name, samplin
         lecture=example["lecture"], 
         solution=example["solution"], 
         prompt_answer_and_solution=prompt_answer_and_solution)
-    
   
-    # default processer
-    # processor = AutoProcessor.from_pretrained("Qwen/Qwen2-VL-2B-Instruct")
-    processor = AutoProcessor.from_pretrained("Qwen/Qwen2-VL-2B-Instruct", min_pixels=256*28*28, max_pixels=512*28*28, padding_side="right")
+    processor = AutoProcessor.from_pretrained(model_name, min_pixels=256*28*28, max_pixels=512*28*28, padding_side="right")
 
-    if image:
-        messages = [
-            {
-                "role": "user",
-                "content": [
-                    {
-                        "type": "image",
-                        "image": image,
-                    },
-                    {"type": "text", "text": text_input},
-                ],
-            }
-        ]
-    else:
-        messages = [
-            {
-                "role": "user",
-                "content": [
-                    {"type": "text", "text": text_input},
-                ],
-            }
-        ]
+    #create empty image if none is given
+    if not image:
+        image = Image.new("RGB", (224, 224), (0, 0, 0))
+        
+    # Simulate chat interaction
+    messages = [
+        {
+            "role": "user",
+            "content": [
+                {
+                    "type": "image",
+                    "image": image,
+                },
+                {"type": "text", "text": text_input},
+            ],
+        }
+    ]
     
[…]

### 2025-02-07 bwcluster/finetuning_scripts/perform_prefix_tuning_experiment.py
bwcluster scripts: training, inference, jobs

@@ -1,0 +1,295 @@
+import argparse
+import gc
+import logging
+import numpy as np
+import pandas as pd
+import torch
+import warnings
+
+from datasets import load_dataset
+from datetime import datetime
+from PIL import Image
+from torch.utils.data import DataLoader
+from tqdm import tqdm
+
+
+from qwen_vl_utils import process_vision_info #this is some basic if-else image processing, can be reused as is for paligemma runs
+from prefix_tuning import PrefixTuningModelPastKeyValues, PrefixDataset, prefix_collate
+from helpers import *
+
+
+"""
+Script to prefix-train a generative MLLM
+Was initially written for qwen, current adaptation is for paligemma
+The training data is the multimodal ScienceQA dataset. 
+The targets are free-text solutions, which are compared to solutions generated by the model. The answer field (int: [1:4]) is not used in this implementation.
+Assumed model architecture: autoregressive textual model + visual encoder
+Script only runs on cuda and does not support CPU-only runs
+"""
+warnings.filterwarnings("ignore", message="You are passing both `text` and `images` to `PaliGemmaProcessor`. The processor expects special image tokens in the text, as many tokens as there are images per each text. It is recommended to add `<image>` tokens in the very beginning of your text and `<bos>` token after that. For this call, we will infer how many images each text has and add special tokens.")
+
+# Empty cache just in case
+torch.cuda.empty_cache()
+
+# Logging
+# Generate timestamped filename
+start_time = datetime.now().strftime("%Y%m%d_%H%M%S")
+log_filename = f"prefix-tuning_{start_time}.log"
+# Configure logging
+logging.basicConfig(
+    level=logging.INFO,
+    format="%(asctime)s - %(levelname)s - %(message)s",
+    handlers=[
+        logging.FileHandler(log_filename),  # Log to file
+        logging.StreamHandler()            # Also log to console (optional)
+    ]
+)
+
+# CONSTANTS
+#NUM_EPOCHS_FT = 20
+#BATCH_SIZE = 4
+#logging.info(f'Set constants: NUM_EPOCHS_FT={NUM_EPOCHS_FT}, BATCH_SIZE={BATCH_SIZE}')
+#logging.info(f'Set constants: NUM_EPOCHS_FT={NUM_EPOCHS_FT}')
+
+device = 'cuda' # torch.cuda.get_device_name(0)
+logging.info(torch.cuda.get_device_name(0))
[…]

### 2025-02-07 bwcluster/finetuning_scripts/perform_prefix_tuning_experiment_dev.py

@@ -1,0 +1,292 @@
+import argparse
+import gc
+import logging
+import numpy as np
+import pandas as pd
+import torch
+import warnings
+
+from datasets import load_dataset
+from datetime import datetime
+from PIL import Image
+from torch.utils.data import DataLoader
+from tqdm import tqdm
+
+
+from qwen_vl_utils import process_vision_info #this is some basic if-else image processing, can be reused as is for paligemma runs
+from prefix_tuning import PrefixTuningModelPastKeyValues, PrefixDataset, prefix_collate
+from helpers import *
+
+
+"""
+Script to prefix-train a generative MLLM
+Was initially written for qwen, current adaptation is for paligemma
+The training data is the multimodal ScienceQA dataset. 
+The targets are free-text solutions, which are compared to solutions generated by the model. The answer field (int: [1:4]) is not used in this implementation.
+Assumed model architecture: autoregressive textual model + visual encoder
+Script only runs on cuda and does not support CPU-only runs
+"""
+warnings.filterwarnings("ignore", message="You are passing both `text` and `images` to `PaliGemmaProcessor`. The processor expects special image tokens in the text, as many tokens as there are images per each text. It is recommended to add `<image>` tokens in the very beginning of your text and `<bos>` token after that. For this call, we will infer how many images each text has and add special tokens.")
+
+# Empty cache just in case
+torch.cuda.empty_cache()
+
+# Logging
+# Generate timestamped filename
+start_time = datetime.now().strftime("%Y%m%d_%H%M%S")
+log_filename = f"prefix-tuning_{start_time}.log"
+# Configure logging
+logging.basicConfig(
+    level=logging.INFO,
+    format="%(asctime)s - %(levelname)s - %(message)s",
+    handlers=[
+        logging.FileHandler(log_filename),  # Log to file
+        logging.StreamHandler()            # Also log to console (optional)
+    ]
+)
+
+# CONSTANTS
+NUM_EPOCHS_FT = 1
+#BATCH_SIZE = 4
+#logging.info(f'Set constants: NUM_EPOCHS_FT={NUM_EPOCHS_FT}, BATCH_SIZE={BATCH_SIZE}')
+logging.info(f'Set constants: NUM_EPOCHS_FT={NUM_EPOCHS_FT}')
+
+device = 'cuda' # torch.cuda.get_device_name(0)
+logging.info(torch.cuda.get_device_name(0))
[…]

### 2025-02-07 bwcluster/finetuning_scripts/perform_prefix_tuning_peft.py

@@ -1,0 +1,394 @@
+import argparse
+import gc
+import logging
+import numpy as np
+import pandas as pd
+import torch
+import warnings
+
+from datasets import load_dataset
+from datetime import datetime
+from peft import PrefixTuningConfig, get_peft_model
+from PIL import Image
+from torch.utils.data import DataLoader
+from tqdm import tqdm
+from trl import SFTConfig
+
+
+from qwen_vl_utils import process_vision_info #this is some basic if-else image processing, can be reused as is for paligemma runs
+from prefix_tuning import PrefixTuningModelPastKeyValues, PrefixDataset, prefix_collate
+from helpers import *
+
+
+"""
+Script to prefix-train a generative MLLM
+Was initially written for qwen, current adaptation is for paligemma
+The training data is the multimodal ScienceQA dataset. 
+The targets are free-text solutions, which are compared to solutions generated by the model. The answer field (int: [1:4]) is not used in this implementation.
+Assumed model architecture: autoregressive textual model + visual encoder
+Script only runs on cuda and does not support CPU-only runs
+"""
+warnings.filterwarnings("ignore", message="You are passing both `text` and `images` to `PaliGemmaProcessor`. The processor expects special image tokens in the text, as many tokens as there are images per each text. It is recommended to add `<image>` tokens in the very beginning of your text and `<bos>` token after that. For this call, we will infer how many images each text has and add special tokens.")
+
+# Empty cache just in case
+torch.cuda.empty_cache()
+
+# Logging
+# Generate timestamped filename
+start_time = datetime.now().strftime("%Y%m%d_%H%M%S")
+log_filename = f"prefix-tuning_{start_time}.log"
+# Configure logging
+logging.basicConfig(
+    level=logging.INFO,
+    format="%(asctime)s - %(levelname)s - %(message)s",
+    handlers=[
+        logging.FileHandler(log_filename),  # Log to file
+        logging.StreamHandler()            # Also log to console (optional)
+    ]
+)
+
+device = 'cuda' # torch.cuda.get_device_name(0)
+logging.info(torch.cuda.get_device_name(0))
+logging.info('Memory Usage:')
+logging.info(f'Allocated: {round(torch.cuda.memory_allocated(0)/1024**3,1)} GB')
+logging.info(f'Cached: {round(torch.cuda.memory_reserved(0)/1024**3,1)} GB')
+
+
+def get_question_text(problem):
+    question = problem['question']
+    return question
+
+
[…]

### 2025-02-07 bwcluster/finetuning_scripts/prefix_tuning.py

@@ -1,0 +1,81 @@
+import torch
+
+from torch import nn
+from torch.utils.data import Dataset
+from transformers import AutoModelForSeq2SeqLM, AutoTokenizer, DynamicCache
+
+# Prefix Tuning Components are defined here
+# Inspiration: https://arxiv.org/pdf/2101.00190
+# Idea: avoid fine-tuning whole (generative) model; instead, learn a prefix for the specific task.
+# Prefix is of fixed length, e.g. 10 tokens, that are learnt from train dataset.
+# Current design decisions: 
+# 1) these tokens are prepended to normal input with the hope of better explaining the taski.
+# 2) the tokens do not directly map to some existing vocabulary items: the additional MLP layer generates embeddings and not input_ids, those embeddings might not correspond to anything meaningful
+# This script directly works with past_key_values
+
+class PrefixTuningPastKeyValues(nn.Module):
+    def __init__(self, match_n_layer, match_n_head, n_embd, device, prefix_length=30):
+        super().__init__()
+        self.device = device
+        self.prefix_length = prefix_length
+        self.input_tokens = torch.arange(prefix_length).long()
+        self.match_n_layer = match_n_layer
+        self.match_n_head = match_n_head
+        self.match_n_embd = n_embd // match_n_head
+        self.n_embd = n_embd
+        #P'_theta
+        self.prefix_embedding = nn.Embedding(self.prefix_length, self.n_embd, dtype=torch.bfloat16)
+        #MLP_theta
+        self.mid_dim = self.n_embd
+        self.mlp = nn.Sequential(
+            nn.Linear(self.n_embd, self.mid_dim, dtype=torch.bfloat16),
+            nn.Tanh(),
+            nn.Linear(self.mid_dim, (self.match_n_layer * 2 * self.n_embd), dtype=torch.bfloat16)
+        )
+
+    def forward(self, bsz):
+        input_tokens = self.input_tokens.unsqueeze(0).expand(bsz, -1).to(self.device)
+        prefix = self.prefix_embedding(input_tokens)
+        past_key_values = self.mlp(prefix)
+        bsz, seqlen, _ = past_key_values.shape
+        past_key_values = past_key_values.view(bsz, seqlen, 2*self.match_n_layer, self.match_n_head, self.match_n_embd)
+        past_key_values = past_key_values.permute([2, 0, 3, 1, 4]).split(2)
+        past_key_values = DynamicCache.from_legacy_cache(past_key_values)
[…]

### 2025-02-07 bwcluster/inference_scripts/inference_paligemma_pref_model.py

@@ -1,0 +1,324 @@
+# Script to download and inference with mistral models
+
+import csv
+import os
+import pandas as pd
+import torch
+
+from collections import defaultdict
+from transformers import Qwen2VLForConditionalGeneration, AutoTokenizer, AutoProcessor
+from qwen_vl_utils import process_vision_info
+
+import argparse
+import logging
+from datetime import datetime
+
+# Generate timestamped filename
+start_time = datetime.now().strftime("%Y%m%d_%H%M%S")
+log_filename = f"inference_qwen_pref_{start_time}.log"
+
+# Configure logging
+logging.basicConfig(
+    level=logging.INFO,
+    format="%(asctime)s - %(levelname)s - %(message)s",
+    handlers=[
+        logging.FileHandler(log_filename),  # Log to file
+        logging.StreamHandler()            # Also log to console (optional)
+    ]
+)
+
+#from vllm import LLM, TokensPrompt
+#from vllm import EngineArgs, LLMEngine
+#from vllm.sampling_params import SamplingParams
+#from vllm.multimodal import MultiModalDataBuiltins
+
+from huggingface_hub import login
+from datasets import load_dataset
+
+from PIL import Image
+
+## UTILS: LOAD MODEL
+# login to hf -- paligemma does not allow unauthorized downloads
+def login_to_hf():
+    try:
+        with open("hf_token", "r") as f:
+            token = f.read().strip()
+    except FileNotFoundError:
+        raise FileNotFoundError("HuggingFace token file not found. Please ensure 'hf_token' exists.")
+        return
+
+    login(token=token)
+
+    logging.info("Login with token: done")
+
+DEVICE = 'cuda'
+
+# download or load from cache the model, put on GPU or distribute among GPUs, prepare CUDA graph
+def prepare_model(model_name, local_adapter_path, checkpoint, max_model_len):
+
+    import torch
+    from peft import PeftModel, PeftConfig
+    #from transformers import AutoModelForCausalLM
+
+    from transformers import AutoModelForImageTextToText
+    model = AutoModelForImageTextToText.from_pretrained(
+            model_name,
+            torch_dtype=torch.bfloat16, #"auto"
+            device_map="auto",
+    ).to(DEVICE)
+    logging.info('Loaded model')
+
+    model.config.vocab_size = model.config._vocab_size
+    model.config.hidden_size = model.config.hidden_size // 2
+
[…]

### 2025-02-07 bwcluster/inference_scripts/inference_qwen_model.py

@@ -1,0 +1,255 @@
+# Script to download and inference with mistral models
+
+import csv
+import os
+import pandas as pd
+from collections import defaultdict
+from transformers import Qwen2VLForConditionalGeneration, AutoTokenizer, AutoProcessor
+from qwen_vl_utils import process_vision_info
+
+import argparse
+import logging
+from datetime import datetime
+
+# Generate timestamped filename
+start_time = datetime.now().strftime("%Y%m%d_%H%M%S")
+log_filename = f"inference_{start_time}.log"
+
+# Configure logging
+logging.basicConfig(
+    level=logging.INFO,
+    format="%(asctime)s - %(levelname)s - %(message)s",
+    handlers=[
+        logging.FileHandler(log_filename),  # Log to file
+        logging.StreamHandler()            # Also log to console (optional)
+    ]
+)
+
+from vllm import LLM, TokensPrompt
+from vllm import EngineArgs, LLMEngine
+from vllm.sampling_params import SamplingParams
+from vllm.multimodal import MultiModalDataBuiltins
+
+from huggingface_hub import login
+from datasets import load_dataset
+
+from PIL import Image
+
+## UTILS: LOAD MODEL
+# login to hf -- mistral does not allow unauthorized downloads
+def login_to_hf():
+    try:
+        with open("hf_token", "r") as f:
+            token = f.read().strip()
+    except FileNotFoundError:
+        raise FileNotFoundError("HuggingFace token file not found. Please ensure 'hf_token' exists.")
+        return
+
+    login(token=token)
+
+    logging.info("Login with token: done")
+
+# download or load from cache the model, put on GPU or distribute among GPUs, prepare CUDA graph
+def prepare_model(model_name, tensor_parallel_size, cpu_offload_gb, max_model_len):
+    # from documentation - default: Load the model on the available device(s)
+    # model = Qwen2VLForConditionalGeneration.from_pretrained(
+    #     "Qwen/Qwen2-VL-2B-Instruct", torch_dtype="auto", device_map="auto"
+    # )
+    llm = LLM(model=model_name, 
+              tokenizer_mode="auto",
+              task="generate",
+              tensor_parallel_size=tensor_parallel_size,  # Can be adjusted for multi-GPU setup
+              cpu_offload_gb=cpu_offload_gb,
+              dtype="float16",
+              # max_model_len=max_model_len 
+             )
+
+    logging.info("Load model: done")
+    
+    return llm
+
+## UTILS: LOAD DATA
[…]

### 2025-02-07 bwcluster/inference_scripts/inference_qwen_pref_model.py

@@ -1,0 +1,316 @@
+# Script to download and inference with mistral models
+
+import csv
+import os
+import pandas as pd
+import torch
+
+from collections import defaultdict
+from transformers import Qwen2VLForConditionalGeneration, AutoTokenizer, AutoProcessor
+from qwen_vl_utils import process_vision_info
+
+import argparse
+import logging
+from datetime import datetime
+
+# Generate timestamped filename
+start_time = datetime.now().strftime("%Y%m%d_%H%M%S")
+log_filename = f"inference_qwen_pref_{start_time}.log"
+
+# Configure logging
+logging.basicConfig(
+    level=logging.INFO,
+    format="%(asctime)s - %(levelname)s - %(message)s",
+    handlers=[
+        logging.FileHandler(log_filename),  # Log to file
+        logging.StreamHandler()            # Also log to console (optional)
+    ]
+)
+
+#from vllm import LLM, TokensPrompt
+#from vllm import EngineArgs, LLMEngine
+#from vllm.sampling_params import SamplingParams
+#from vllm.multimodal import MultiModalDataBuiltins
+
+#from huggingface_hub import login
+from datasets import load_dataset
+
+from PIL import Image
+
+## UTILS: LOAD MODEL
+# login to hf -- mistral does not allow unauthorized downloads
+#def login_to_hf():
+#    try:
+#        with open("hf_token", "r") as f:
+#            token = f.read().strip()
+#    except FileNotFoundError:
+#        raise FileNotFoundError("HuggingFace token file not found. Please ensure 'hf_token' exists.")
+#        return
+
+#    login(token=token)
+
+#    logging.info("Login with token: done")
+
+DEVICE = 'cuda'
+
+# download or load from cache the model, put on GPU or distribute among GPUs, prepare CUDA graph
+def prepare_model(model_name, local_adapter_path, checkpoint, tensor_parallel_size, cpu_offload_gb, max_model_len):
+
+    import torch
+    from peft import PeftModel, PeftConfig
+    #from transformers import AutoModelForCausalLM
+
+    from transformers import AutoModelForImageTextToText
+    model = AutoModelForImageTextToText.from_pretrained(
+            "Qwen/Qwen2-VL-2B-Instruct",
+            torch_dtype=torch.bfloat16, #"auto"
+            device_map="auto",
+    ).to(DEVICE)
+    logging.info('Loaded model')
+
[…]

### 2025-02-07 bwcluster/jobs/inference_model_QTCHL.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --job-name=qtchlSciQA
+#SBATCH --output=job_output_qtchl_%j.log
+#SBATCH --gres=gpu:1
+#SBATCH --time=01:30:00
+#SBATCH --partition=gpu_4,gpu_4_h100
+export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
+source ~/fmenv/bin/activate
+python inference_paligemma_model.py --ngpus 1 --setting QTCHL --resume 2400 --split validation --model_name "google/paligemma-3b-ft-science-qa-224"

### 2025-02-07 bwcluster/jobs/inference_model_QTCHLS.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --job-name=qtchlsSciQA
+#SBATCH --output=job_output_qtchls_%j.log
+#SBATCH --gres=gpu:1
+#SBATCH --time=02:00:00
+#SBATCH --partition=gpu_4,gpu_4_h100
+export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
+source ~/fmenv/bin/activate
+python inference_paligemma_model.py --ngpus 1 --setting QTCHLS --resume 2375 --split validation --model_name "google/paligemma-3b-ft-science-qa-224"

### 2025-02-07 bwcluster/jobs/inference_model_QTCHLS_test.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --job-name=qtchlsSciQAtest
+#SBATCH --output=job_output_qtchls_test_%j.log
+#SBATCH --gres=gpu:1
+#SBATCH --time=01:30:00
+#SBATCH --partition=gpu_4,gpu_4_h100
+export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
+source ~/fmenv/bin/activate
+python inference_paligemma_model.py --ngpus 1 --setting QTCHLS --resume 2396 --split test --model_name "google/paligemma-3b-ft-science-qa-224"

### 2025-02-07 bwcluster/jobs/inference_model_QTCHL_test.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --job-name=qtchlSciQAtest
+#SBATCH --output=job_output_qtchl_test_%j.log
+#SBATCH --gres=gpu:1
+#SBATCH --time=02:30:00
+#SBATCH --partition=gpu_4,gpu_4_h100
+export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
+source ~/fmenv/bin/activate
+python inference_paligemma_model.py --ngpus 1 --setting QTCHL --resume 2262 --split test --model_name "google/paligemma-3b-ft-science-qa-224"

### 2025-02-07 bwcluster/jobs/inference_model_QTCHS.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --job-name=qtchsSciQA
+#SBATCH --output=job_output_qtchs_%j.log
+#SBATCH --gres=gpu:1
+#SBATCH --time=02:00:00
+#SBATCH --partition=gpu_4,gpu_4_h100
+export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
+source ~/fmenv/bin/activate
+python inference_paligemma_model.py --ngpus 1 --setting QTCHS --resume 2378 --split validation --model_name "google/paligemma-3b-ft-science-qa-224"

### 2025-02-07 bwcluster/jobs/inference_model_QTCHS_test.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --job-name=qtchsSciQAtest
+#SBATCH --output=job_output_qtchs_test_%j.log
+#SBATCH --gres=gpu:1
+#SBATCH --time=01:30:00
+#SBATCH --partition=gpu_4,gpu_4_h100
+export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
+source ~/fmenv/bin/activate
+python inference_paligemma_model.py --ngpus 1 --setting QTCHS --resume 2395 --split test --model_name "google/paligemma-3b-ft-science-qa-224"

### 2025-02-07 bwcluster/jobs/inference_model_QTCH_test.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --job-name=qtchQwenPtTest
+#SBATCH --output=job_output_qtch_qwen_test_%j.log
+#SBATCH --gres=gpu:1
+#SBATCH --time=02:30:00
+#SBATCH --partition=gpu_4,gpu_4_h100,gpu_8,gpu_4_h100
+export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
+source ~/fmenv/bin/activate
+python inference_qwen_model.py --ngpus 1 --setting QTCH --resume 0 --split test --model_name "Qwen/Qwen2-VL-2B-Instruct"

### 2025-02-07 bwcluster/jobs/inference_model_cpu_QTCH.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --job-name=CPUqtchSciQA
+#SBATCH --output=job_output_qtch_cpu_%j.log
+#SBATCH --time=00:15:00
+#SBATCH --nodes=2
+#SBATCH --partition=dev_multiple
+export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
+source ~/fmenv/bin/activate
+python inference_qwen_model.py --ngpus=0  --setting QTCH --resume 0

### 2025-02-07 bwcluster/jobs/inference_peft_paligemma_golden_QTCH_test.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --job-name=qtchPlgPEFT
+#SBATCH --output=job_output_qtch_pref_paligemma_golden_test_%j.log
+#SBATCH --gres=gpu:1
+#SBATCH --time=03:30:00
+#SBATCH --partition=gpu_4,gpu_8,gpu_4_h100,gpu_4_a100
+export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
+source ~/fmenv/bin/activate
+python inference_paligemma_pref_model.py --setting QTCH --resume 0 --split test --local_path "google_paligemma2-3b-pt-224_golden" --checkpoint 57580

### 2025-02-07 bwcluster/jobs/inference_peft_paligemma_teacher_QTCH_test.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --job-name=qtchPEFTplgKDtest
+#SBATCH --output=job_output_qtch_peft_paligemma_teacher_test_%j.log
+#SBATCH --gres=gpu:1
+#SBATCH --time=03:30:00
+#SBATCH --partition=gpu_4,gpu_8,gpu_4_h100,gpu_4_a100
+export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
+source ~/fmenv/bin/activate
+python inference_paligemma_pref_model.py --setting QTCH --resume 0 --split test --local_path "google_paligemma2-3b-pt-224_teacher" --checkpoint 57580

### 2025-02-07 bwcluster/jobs/inference_peft_paligemma_teacher_QTCH_test_dev.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --job-name=qtchPEFTplgKDtest
+#SBATCH --output=job_output_qtch_peft_paligemma_teacher_test_%j.log
+#SBATCH --gres=gpu:1
+#SBATCH --time=00:30:00
+#SBATCH --partition=dev_gpu_4,gpu_8,gpu_4_h100,dev_gpu_4_a100
+export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
+source ~/fmenv/bin/activate
+python inference_paligemma_pref_model.py --setting QTCH --resume 0 --split test --local_path "google_paligemma2-3b-pt-224_teacher" --checkpoint 57580

### 2025-02-07 bwcluster/jobs/inference_peft_qwen_golden_QTCH_test.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --job-name=qtchPEFTqwnG
+#SBATCH --output=job_output_peft_qwen_golden_test_%j.log
+#SBATCH --gres=gpu:1
+#SBATCH --time=06:30:00
+#SBATCH --partition=gpu_4,gpu_8,gpu_4_h100,gpu_4_a100
+export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
+source ~/fmenv/bin/activate
+python inference_qwen_pref_model.py --ngpus 1 --setting QTCH --resume 0 --split test --local_path "qwen_qwen2-vl-2b-instruct_golden" --checkpoint 57580

### 2025-02-07 bwcluster/jobs/inference_peft_qwen_teacher_QTCH_test.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --job-name=qtchPEFTqwenKD
+#SBATCH --output=job_output_qtch_peft_qwen_teacher_test_%j.log
+#SBATCH --gres=gpu:1
+#SBATCH --time=06:30:00
+#SBATCH --partition=gpu_4,gpu_8
+export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
+source ~/fmenv/bin/activate
+python inference_qwen_pref_model.py --ngpus 1 --setting QTCH --resume 0 --split test --local_path "qwen_qwen2-vl-2b-instruct_teacher" --checkpoint 57580

### 2025-02-07 bwcluster/jobs/inference_qwen_model_upd.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --job-name=qtchQwen
+#SBATCH --output=job_output_qtch_qwen_test_%j.log
+#SBATCH --gres=gpu:1
+#SBATCH --time=01:30:00
+#SBATCH --partition=gpu_4,gpu_8,gpu_4_h100,gpu_4_a100
+export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
+source ~/fmenv/bin/activate
+python inference_qwen_model2.py --setting QTCH --resume 0 --split test --model_name "Qwen/Qwen2-VL-2B-Instruct"

### 2025-02-07 bwcluster/jobs/inference_raw_model_QTCH.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --job-name=RAWqtchSciQA
+#SBATCH --output=job_output_qtch_raw_%j.log
+#SBATCH --gres=gpu:1
+#SBATCH --time=01:00:00
+#SBATCH --partition=gpu_4_h100
+export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
+source ~/fmenv/bin/activate
+python inference_paligemma_model.py --ngpus 1 --setting QTCH --resume 2315 --split validation

### 2025-02-07 bwcluster/jobs/inference_raw_model_QTCHL.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --job-name=RAWqtchlSciQA
+#SBATCH --output=job_output_qtchl_raw_%j.log
+#SBATCH --gres=gpu:1
+#SBATCH --time=00:30:00
+#SBATCH --partition=gpu_4,gpu_4_h100
+export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
+source ~/fmenv/bin/activate
+python inference_paligemma_model.py --ngpus 1 --setting QTCHL --resume 3071 --split validation

### 2025-02-07 bwcluster/jobs/inference_raw_model_QTCHLS.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --job-name=RAWqtchlsSciQA
+#SBATCH --output=job_output_qtchls_raw_%j.log
+#SBATCH --gres=gpu:1
+#SBATCH --time=00:30:00
+#SBATCH --partition=gpu_4,gpu_4_h100
+export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
+source ~/fmenv/bin/activate
+python inference_paligemma_model.py --ngpus 1 --setting QTCHLS --resume 3194 --split validation

### 2025-02-07 bwcluster/jobs/inference_raw_model_QTCHLS_test.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --job-name=RAWqtchlsSciQAtest
+#SBATCH --output=job_output_qtchls_test_raw_%j.log
+#SBATCH --gres=gpu:1
+#SBATCH --time=00:30:00
+#SBATCH --partition=dev_gpu_4_a100,dev_gpu_4
+export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
+source ~/fmenv/bin/activate
+python inference_paligemma_model.py --ngpus 1 --setting QTCHLS --resume 3065 --split test

### 2025-02-07 bwcluster/jobs/inference_raw_model_QTCHL_test.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --job-name=RAWqtchlSciQAtest
+#SBATCH --output=job_output_qtchl_raw_test_%j.log
+#SBATCH --gres=gpu:2
+#SBATCH --time=00:30:00
+#SBATCH --partition=dev_gpu_4,gpu_4,gpu_4_h100,gpu_4_a100
+export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
+source ~/fmenv/bin/activate
+python inference_paligemma_model.py --ngpus 2 --setting QTCHL --resume 2885 --split test

### 2025-02-07 bwcluster/jobs/inference_raw_model_QTCHS.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --job-name=RAWqtchsSciQA
+#SBATCH --output=job_output_qtchs_raw_%j.log
+#SBATCH --gres=gpu:1
+#SBATCH --time=01:30:00
+#SBATCH --partition=gpu_4_h100
+export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
+source ~/fmenv/bin/activate
+python inference_paligemma_model.py --ngpus 1 --setting QTCHS --resume 2328 --split validation

### 2025-02-07 bwcluster/jobs/inference_raw_model_QTCHS_test.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --job-name=RAWqtchsSciQAtest
+#SBATCH --output=job_output_qtchs_test_raw_%j.log
+#SBATCH --gres=gpu:1
+#SBATCH --time=00:30:00
+#SBATCH --partition=dev_gpu_4,dev_gpu_4_a100
+export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
+source ~/fmenv/bin/activate
+python inference_paligemma_model.py --ngpus 1 --setting QTCHS --resume 2650 --split test

### 2025-02-07 bwcluster/jobs/inference_raw_model_QTCH_test.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --job-name=RAWqtchSciQAtest
+#SBATCH --output=job_output_qtch_raw_test_%j.log
+#SBATCH --gres=gpu:1
+#SBATCH --time=04:00:00
+#SBATCH --partition=gpu_4_a100,gpu_4,gpu_4_h100
+export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
+source ~/fmenv/bin/activate
+python inference_paligemma_model.py --ngpus 1 --setting QTCH --resume 620 --split test

### 2025-02-07 bwcluster/jobs/train_lora.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --job-name=LoRATuneExp
+#SBATCH --output=job_output_lora_%j.log
+#SBATCH --gres=gpu:2
+#SBATCH --time=02:00:00
+#SBATCH --partition=gpu_4,gpu_4_h100,gpu_4_a100
+export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
+source ~/fmenv/bin/activate
+python perform_lora_experiment.py

### 2025-02-07 bwcluster/jobs/train_prefix_KD2.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --job-name=Pref2KD
+#SBATCH --output=job_output_pref_kd2_%j.log
+#SBATCH --gres=gpu:1
+#SBATCH --time=16:00:00
+#SBATCH --partition=gpu_4,gpu_8,gpu_4_h100
+export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
+source ~/fmenv/bin/activate
+python perform_prefix_tuning_experiment.py --exp_setting teacher --batch_size 2

### 2025-02-07 bwcluster/jobs/train_prefix_Param.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --job-name=PrefTuneExp
+#SBATCH --output=job_output_pref_%j.log
+#SBATCH --gres=gpu:1
+#SBATCH --time=02:30:00
+#SBATCH --partition=gpu_4,gpu_4_h100,gpu_4_a100
+export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
+source ~/fmenv/bin/activate
+python perform_prefix_tuning_Param_experiment.py

### 2025-02-07 bwcluster/jobs/train_prefix_dev.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --job-name=PrefDevA100
+#SBATCH --output=job_output_pref_dev_a100_%j.log
+#SBATCH --gres=gpu:1
+#SBATCH --time=00:30:00
+#SBATCH --partition=dev_gpu_4_a100
+export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
+source ~/fmenv/bin/activate
+python perform_prefix_tuning_experiment.py --batch_size 8 --num_epochs 1

### 2025-02-07 bwcluster/jobs/train_prefix_dev2.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --job-name=PrefDev2Exp
+#SBATCH --output=job_output_pref_dev2_%j.log
+#SBATCH --gres=gpu:1
+#SBATCH --time=00:30:00
+#SBATCH --partition=gpu_4,gpu_4_a100,gpu_4_h100,gpu_8,dev_gpu_4,dev_gpu_4_a100
+export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
+source ~/fmenv/bin/activate
+python perform_prefix_tuning_experiment.py --ngpus 1 --batch_size 2

### 2025-02-07 bwcluster/jobs/train_prefix_paligemma.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --job-name=PrfPlgPKVG
+#SBATCH --output=job_output_pref_paligemma_pkv_golden_%j.log
+#SBATCH --gres=gpu:1
+#SBATCH --time=12:00:00
+#SBATCH --partition=gpu_4_h100,gpu_4_a100
+export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
+source ~/fmenv/bin/activate
+python perform_prefix_tuning_experiment.py --model_name "google/paligemma2-3b-pt-224" --exp_setting golden --batch_size 4 --num_epochs 20

### 2025-02-07 bwcluster/jobs/train_prefix_paligemma2.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --job-name=Pref2Plgm
+#SBATCH --output=job_output_pref_paligemma_%j.log
+#SBATCH --gres=gpu:1
+#SBATCH --time=18:00:00
+#SBATCH --partition=gpu_8,gpu_4,gpu_4_h100
+export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
+source ~/fmenv/bin/activate
+python perform_prefix_tuning_experiment.py --ngpus 1 --model_name "google/paligemma2-3b-pt-224" --batch_size 2

### 2025-02-07 bwcluster/jobs/train_prefix_paligemma_KD.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --job-name=PrfKDPlgPKV
+#SBATCH --output=job_output_pref_KD_paligemma_pkv_%j.log
+#SBATCH --gres=gpu:1
+#SBATCH --time=12:00:00
+#SBATCH --partition=gpu_4_h100,gpu_4_a100
+export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
+source ~/fmenv/bin/activate
+python perform_prefix_tuning_experiment.py --model_name "google/paligemma2-3b-pt-224" --exp_setting teacher --batch_size 4 --num_epochs 20

### 2025-02-07 bwcluster/jobs/train_prefix_paligemma_KD2.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --job-name=Pref2KDPaligemma
+#SBATCH --output=job_output_pref_KD_2_2_paligemma_%j.log
+#SBATCH --gres=gpu:1
+#SBATCH --time=18:00:00
+#SBATCH --partition=gpu_4,gpu_8,gpu_4_h100
+export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
+source ~/fmenv/bin/activate
+python perform_prefix_tuning_experiment.py --model_name "google/paligemma2-3b-pt-224" --exp_setting teacher --batch_size 2

### 2025-02-07 bwcluster/jobs/train_prefix_peft_dev.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --job-name=PrefPEFT
+#SBATCH --output=job_output_pref_PEFT_%j.log
+#SBATCH --gres=gpu:1
+#SBATCH --time=00:15:00
+#SBATCH --partition=dev_gpu_4_a100,gpu_4_a100,gpu_4_h100
+export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
+source ~/fmenv/bin/activate
+python perform_prefix_tuning_peft.py --num_epochs 1 --batch_size 8

### 2025-02-07 bwcluster/jobs/train_prefix_peft_paligemma.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --job-name=PEFTplgG
+#SBATCH --output=job_output_pref_paligemma_PEFT_golden_%j.log
+#SBATCH --gres=gpu:1
+#SBATCH --time=10:00:00
+#SBATCH --partition=gpu_4_a100,gpu_4_h100
+export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
+source ~/fmenv/bin/activate
+python perform_prefix_tuning_peft.py --num_epochs 20 --model_name "google/paligemma2-3b-pt-224" --exp_setting golden

### 2025-02-07 bwcluster/jobs/train_prefix_peft_paligemma_KD.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --job-name=PEFTPlgKD
+#SBATCH --output=job_output_pref_PEFT_plg_KD_%j.log
+#SBATCH --gres=gpu:1
+#SBATCH --time=10:00:00
+#SBATCH --partition=gpu_4_a100,gpu_4_h100
+export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
+source ~/fmenv/bin/activate
+python perform_prefix_tuning_peft.py --exp_setting teacher --model_name "google/paligemma2-3b-pt-224" --num_epochs 20

### 2025-02-07 bwcluster/jobs/train_prefix_peft_paligemma_dev.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --job-name=PrfPlgPEFT
+#SBATCH --output=job_output_pref_PEFT_plg_dev_%j.log
+#SBATCH --gres=gpu:1
+#SBATCH --time=00:30:00
+#SBATCH --partition=gpu_4_a100,gpu_4_h100,dev_gpu_4_a100,dev_gpu_4
+export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
+source ~/fmenv/bin/activate
+python perform_prefix_tuning_peft.py --model_name "google/paligemma2-3b-pt-224" --batch_size 2 --num_epochs 2

### 2025-02-07 bwcluster/test_reload_from_checkpoint.py

@@ -1,0 +1,24 @@
+import torch
+from peft import PeftModel, PeftConfig
+#from transformers import AutoModelForCausalLM
+
+from transformers import AutoModelForImageTextToText, Qwen2VLForConditionalGeneration
+print('Hi there!')
+model = AutoModelForImageTextToText.from_pretrained(
+            "Qwen/Qwen2-VL-2B-Instruct",
+            torch_dtype=torch.bfloat16, #"auto"
+            device_map="auto",
+)
+print('Loaded model')
+
+config = PeftConfig.from_pretrained("qwen_qwen2-vl-2b-instruct_golden/checkpoint-9500") #checkpoint-9500
+print('Loaded config')
+print(config)
+#model = AutoModelForCausalLM.from_pretrained("google/gemma-2b")
+model = PeftModel.from_pretrained(model, 
+    "qwen_qwen2-vl-2b-instruct_golden/checkpoint-9500",
+    #is_trainable=True # 👈 here
+    )
+# check if it's working
+#model.print_trainable_parameters()
+#model()

### 2025-02-07 adapter_experiments/perform_prefix_tuning_experiment.py
reorganisation and cleanup

@@ -11,7 +11,6 @@ from PIL import Image
 from torch.utils.data import DataLoader
 from tqdm import tqdm
 
-from transformers import AutoModelForImageTextToText, AutoTokenizer, AutoProcessor, Qwen2VLForConditionalGeneration
 
 from qwen_vl_utils import process_vision_info #this is some basic if-else image processing, can be reused as is for paligemma runs
 from prefix_tuning import PrefixTuningModelPastKeyValues, PrefixDataset, prefix_collate
@@ -46,8 +45,9 @@ logging.basicConfig(
 
 # CONSTANTS
 NUM_EPOCHS_FT = 20
-BATCH_SIZE = 4
-logging.info(f'Set constants: NUM_EPOCHS_FT={NUM_EPOCHS_FT}, BATCH_SIZE={BATCH_SIZE}')
+#BATCH_SIZE = 4
+#logging.info(f'Set constants: NUM_EPOCHS_FT={NUM_EPOCHS_FT}, BATCH_SIZE={BATCH_SIZE}')
+logging.info(f'Set constants: NUM_EPOCHS_FT={NUM_EPOCHS_FT}')
 
 device = 'cuda' # torch.cuda.get_device_name(0)
 logging.info(torch.cuda.get_device_name(0))
@@ -65,10 +65,22 @@ def preprocess_input_qwen(tokenizer, processor, prompts, texts, images, y, devic
         text=messages,
         images=image_inputs,
         videos=video_inputs,
-        padding=True,
+        padding="longest",
         return_tensors="pt",
     )
     max_length = inputs["input_ids"].size(1) 
+    labels = tokenizer(y, padding="max_length", truncation=True, max_length=max_length, return_tensors="pt")["input_ids"]
+    return inputs.to(device, dtype=torch.bfloat16), labels.to(device)
+
+def preprocess_input_paligemma(tokenizer, processor, prompts, texts, images, y, device):
+    images = [image.resize((224, 224)) for image in images]
+    inputs = processor(
+        text=prompts,
+        images=images,
+        return_tensors="pt",
+        padding="longest"
+    )
+    max_length = inputs["input_ids"].size(1)
     labels = tokenizer(y, padding="max_length", truncation=True, max_length=max_length, return_tensors="pt")["input_ids"]
     return inputs.to(device, dtype=torch.bfloat16), labels.to(device)
 
@@ -108,7 +120,8 @@ def train(model, tokenizer, processor, optimizer, dataloader_train, dataloader_v
[…]

### 2025-02-07 dataset_utils.py

@@ -1,55 +1,0 @@
-# Dataset-specific utils. TODO: separate file
-def get_question_text(problem):
-    question = problem['question']
-    return question
-
-
-def get_choice_text(probelm, options):
-    choices = probelm['choices']
-    choice_list = []
-    for i, c in enumerate(choices):
-        choice_list.append("({}) {}".format(options[i], c))
-    choice_txt = " ".join(choice_list)
-    return choice_txt
-
-
-def get_context_text(problem, use_caption):
-    txt_context = problem['hint']
-    img_context = problem['caption'] if use_caption else ""
-    context = " ".join([txt_context, img_context]).strip()
-    if context == "":
-        context = "N/A"
-    return context
-
-def build_prompt(question_data, use_lecture=False, use_solution=False):
-    question = get_question_text(question_data)
-    choices = get_choice_text(question_data, [choice_num for choice_num in range(5)])
-    hint = get_context_text(question_data, False)
-    #image = question_data['image']
-    task = question_data['task']
-    input_prompt = f'Question: {question}\n Task: {task}\n Choices: {choices}\n Hint: {hint}'
-    if use_lecture:
-        lecture = f'\n Lecture: {question_data["lecture"]}'
-        input_prompt += lecture
-    if use_solution and question_data["solution"]:
-        solution = f'\n Solution: {question_data["solution"]}'
-        input_prompt += solution
-    prompt = [input_prompt]
-    #if image:
-    #    prompt.append(image)
-    return prompt
-
-def build_message(row):
-    messages = [
-        {
-            "role": "user",
-            "content": [
-                {
-                    "type": "image",
-                    "image": row["image"],
-                },
-                {"type": "text", "text": row['input']},
-            ],
-        }
-    ]
-    return messages

### 2025-02-07 train_model.sh

@@ -1,9 +1,0 @@
-#!/bin/bash
-#SBATCH --job-name=PrefTuneExp
-#SBATCH --output=job_output_pref_%j.log
-#SBATCH --gres=gpu:1
-#SBATCH --time=01:30:00
-#SBATCH --partition=gpu_4,gpu_4_h100,gpu_4_a100
-export PYTHONPATH=/pfs/data5/home/st/st_us-053000/st_st186032/fmenv/lib/python3.9/site-packages
-source ~/fmenv/bin/activate
-python perform_tuning_experiment.py

### 2025-02-07 adapter_experiments/perform_tuning_experiment.py
bug fix perform_tuning_experiment.py
the script was apparently an old version with a bug in saving train and val errors

@@ -181,10 +181,10 @@ logging.info('Saving assets')
 saving_path = f"{model_name.replace('/', '_').lower()}_pref_{NUM_EPOCHS_FT}_{BATCH_SIZE}"
 torch.save(model.state_dict(), saving_path)
 logging.info(f'Saved model to path: {saving_path}')
-with open(f'saving_path_{train_error}', 'w') as f:
+with open(f'{saving_path}_train_error', 'w') as f:
     f.write(train_errors_ft_qwen)
     logging.info('Saved train_errors')
-with open(f'saving_path_{val_error}', 'w') as f:
+with open(f'{saving_path}_val_error', 'w') as f:
     f.write(val_errors_ft_qwen)
     logging.info('Saved validation_errors')
 

### 2025-02-07 adapter_experiments/perform_tuning_experiment.py
rename dependency perform_tuning_experiment.py
the script was also referring to a copy of the helpers.py called dataset_utils.py

@@ -11,7 +11,7 @@ from tqdm import tqdm
 
 from qwen_vl_utils import process_vision_info #this is some basic if-else image processing, can be reused as is for paligemma runs
 from prefix_tuning import PrefixTuningModel, PrefixDataset, prefix_collate
-from dataset_utils import *
+from helpers import *
 
 
 """
